```{r, include=FALSE, message=FALSE}
if (!exists("PREAMBLE_LOADED")) source("preamble.R")


# Store parameters for model fitting
d.mean_sd_scaling <-
  d.for_analysis %>%
  group_by(Phase) %>%
  filter(Item.Labeled == FALSE) %>%
  summarise(across(Item.VOT, list(mean = mean, sd = sd)))

VOT.mean_exposure <- (d.mean_sd_scaling %>% pull(Item.VOT_mean))[1]
VOT.sd_exposure <- (d.mean_sd_scaling %>% pull(Item.VOT_sd))[1]
VOT.mean_test <- (d.mean_sd_scaling %>% pull(Item.VOT_mean))[2]
VOT.sd_test <- (d.mean_sd_scaling %>% pull(Item.VOT_sd))[2]
rm(d.mean_sd_scaling)

levels_Condition.Exposure = c("Shift0", "Shift10", "Shift40")
contrast_type <- "difference"
```

# Results {#sec:results}
We begin by describing our analysis approach. Since this analysis approach is still uncommon in research on speech perception, we demonstrate that it replicates previous findings when applied to our data at the level of analysis employed in previous work: the detection of overall changes in listeners' categorization behavior after exposure. Following this, we turn to the questions of primary interest: *incremental* changes in participants' categorization responses from pre-exposure onward, depending on the type (exposure condition) and amount of exposure (test block). These latter tests allow us to assess predictions (1) and (2a,b) about the role of prior and recent experience in explaining incremental adaptive speech perception, as well as prediction (4) about the diminishing rate of behavioral changes with increasing exposure. To facilitate the interpretation of our results, we introduce normative models (ideal observers) that determine the expected categorization functions of idealized listeners prior to, and following, exposure (following the same approach used in Figure \@ref(fig:predictions)). This allows us to identify previously unrecognized constraints on adaptive speech perception (prediction 3 - learning to convergence).

```{r model-fit-test-blocks, include=FALSE}
fit_test <-
  fit_model(
    data = d.for_analysis,
    phase = "test",
    formulation = "standard",
    priorSD = 15,
    adapt_delta = .995)

fit_exposure <-
  fit_model(
    data = d.for_analysis,
    phase = "exposure",
    formulation = "standard",
    priorSD = 15,
    adapt_delta = .999)
```

## Analysis approach
We analyzed participants' categorization responses during exposure and test blocks in two separate Bayesian mixed-effects psychometric models, using \texttt{brms} [@R-brms_a] in \texttt{R} [@R; @RStudio].^[Fitting the models separately side-steps concerns about how differences in the VOT distribution during exposure blocks might affect the analysis of test blocks. For the test analyses, it also removes any potential collinearity between effects of exposure and effects of VOT. The SI reports additional analyses over the combined data, including extensions of the psychometric models to include lapse rates that can vary by block (\@ref(sec:analysis-lapse)) and non-parametric smooths to model non-linear effects of VOT and exposure (\@ref(sec:GAMM)). All analyses replicate the findings reported here.] Psychometric models account for attentional lapses while estimating participants' categorization functions. Failing to account for attentional lapses---while still common in research on speech perception [for exceptions, see @clayards2008; @kleinschmidt-jaeger2016]---can lead to biased estimates of categorization boundaries [@prins2011; @wichmann-hill2001]. For the advantages of Bayesian psychometric models, we refer to @kuss2005 and @prins2019. For the present experiment, lapse rates were negligible (`r print_CI(fit_test, "theta1_Intercept")`), and all results replicate in simple mixed-effects logistic regressions [@jaeger2008]. This lapse rate compares favorably against those assumed or reported in prior work [e.g., @clayards2008; @kleinschmidt-jaeger2016; @kleinschmidt2020].

The psychometric models for exposure and test blocks each regressed participants' categorization responses against the full factorial interaction of VOT, block, and exposure condition, along with the maximal random effect structure (by-participant intercepts and slopes for VOT, block, and their interaction, and by-item intercept and slopes for the full factorial design; see SI, \@ref(sec:analysis-approach)). All hypothesis tests reported below are based on these models. Figure \@ref(fig:plot-fit-PSE) summarizes the results that we describe in more detail next. Panels A and B show participants’ categorization responses during exposure and test blocks, along with the categorization function estimated from those responses via the mixed-effects psychometric models. These panels facilitate comparison between exposure conditions within each block. Panel C summarizes changes in listeners' point of subject equality (PSE)---i.e., the point along the VOT continuum at which participants are equally likely to respond "d" or "t"---across blocks and conditions. This highlights how the type and amount of phonetic input affect listeners' categorization functions. Here we focus on the test blocks, which were identical within and across exposure conditions. Analyses of the exposure blocks, reported in the SI (\@ref(sec:SI-exposure-block-analysis)), replicate all effects found in the test blocks.

```{r fit-nested-models-for-intercepts-slopes-plot, message=FALSE}
# Refit the same model, formulated as nesting intercepts and slopes within each unique combination of
# exposure condition and block. This makes it easier to extract the intercept, slope, and PSE for the
# large result figure (Panels C and D).
fit_test_nested_within_condition_and_block <-
  fit_model(
    data = d.for_analysis,
    phase = "test",
    formulation = "nested_slope",
    priorSD = 15,
    adapt_delta = .995)

fit_exposure_nested <-
  fit_model(
    data = d.for_analysis,
    phase = "exposure",
    formulation = "nested_slope",
    priorSD = 15,
    adapt_delta = .999)
```

```{r extract-intercepts-slopes-from-test-and-exposure-models}
d.psychometric_intercept.slope.PSE <-
  full_join(
    fit_test_nested_within_condition_and_block %>% get_intercepts_and_slopes(),
    fit_exposure_nested %>% get_intercepts_and_slopes()) %>%
  mutate(
    Block = factor(Block),
    # Adding unscaled slope here to have a more intuitive scale on which to compare VOT slopes
    # It is also one way to put the slopes for both exposure and test on the same scale (since
    # the SDs for exposure and test differ).
    slope_unscaled = slope / (2 * ifelse(Block %in% c(2, 4, 6), VOT.sd_exposure, VOT.sd_test)),
    PSE =
      descale(
        -Intercept/slope,
        VOT.mean_test,
        ifelse(Block %in% c(2, 4, 6), VOT.sd_exposure, VOT.sd_test))) %>%
  group_by(Condition.Exposure, Block) %>%
  median_hdci() %>% 
  add_block_labels()
```

```{r}
# divide database into k-folds
set.seed(2000)
d.chodroff_wilson.isolated %<>%
  group_by(Talker) %>%
  vfold_cv(v = 5)

# get the folds from each train-test split
# nested within each row
d.chodroff_wilson.isolated <- 
  map(
    .x = 1:5, 
    ~ assessment(d.chodroff_wilson.isolated$splits[[.x]]) %>% mutate(fold = .x)) %>%
  list_rbind()

# Make ideal observers that are fit on the exposure VOTs of the three exposure conditions.
# These ideal observers can be thought of as reflecting a learner that has fully learned
# the exposure distributions. Note that we include perceptual noise in the ideal observers
# (estimated in Kronrod et al., 2016) to make their perception more human-like.
io <-    
  make_VOT_IOs_from_exposure(
    d.for_analysis %>%
      filter(Phase == "exposure") %>%
      group_by(ParticipantID, Condition.Exposure) %>%
      nest(data = -c(ParticipantID, Condition.Exposure)) %>%
      group_by(Condition.Exposure) %>%
      # Subset data to a single participant per exposure condition. This is sufficient since
      # the ideal observer's /d/ and /t/ categories only depend on the sufficient statistics
      # (mean and SD) at the end of the experiment, which were identical across participants
      # in each condition.
      #
      # Note that this approach trains the ideal observers on the *actual* VOT distribution
      # that participants heard, not on the theoretical distributions these VOTs were sampled
      # from. This is in line with our goal to simulate behavior of an idealized participant
      # who has fully learned the exposure distributions.
      slice_sample(n = 1) %>%
      unnest(data) %>%
      mutate(category = ifelse(Item.ExpectedResponse.Voicing == "voiced", "/d/", "/t/"))) %>%
      # only VOT is used for categorising stimuli
  left_join(d.for_analysis %>%  
        filter(Phase == "test") %>%
        distinct(Condition.Exposure, Item.VOT) %>%
  rename(x = Item.VOT) %>%
    nest(x = x)) %>%
  # Add prior based on Chodroff & Wilson (2018). Note that this database has substantially higher
  # VOT variance than our exposure conditions, which we based on estimates from Kronrod et al.
  # (2016). Uses all three cues; duration and spectral noise taken from Kronrod et al.
  bind_rows(
    # make sure the IOs to be constructed from database include both folded data and unfolded (all) data  
    bind_rows(
      d.chodroff_wilson.isolated %>%
        group_by(fold) %>%
        nest(data = -fold),
      d.chodroff_wilson.isolated %>%
        ungroup() %>%
        nest(data = everything()) %>%
        mutate(fold = NA)) %>%
      mutate(
        io = map(
          data, ~
            make_MVG_ideal_observer_from_data(
              data = .x,
              cues = c("VOT", "f0_Mel", "vowel_duration"),
              Sigma_noise = matrix(
                c(80, 0, 0, 0, 878, 0, 0, 0, 80),
                nrow = 3,
                dimnames = list(c("VOT", "f0_Mel", "vowel_duration"), c("VOT", "f0_Mel", "vowel_duration")))))) %>%
      mutate(Condition.Exposure = "prior") %>%
      select(-data) %>%
      left_join(
        d.for_analysis %>%  
          filter(Phase == "test") %>%
          distinct(Condition.Exposure, Item.VOT, Item.f0_Mel, Item.VowelDuration) %>%
          mutate(x = pmap(.l = list(Item.VOT, Item.f0_Mel, Item.VowelDuration), ~ c(..1, ..2, ..3))) %>%
          select(Condition.Exposure, x) %>%
          mutate(Condition.Exposure = "prior") %>%
          nest(x = x)) %>%
      mutate(Condition.Exposure = ifelse(is.na(fold), "prior", str_c("prior", fold))))

# Get logistic parameter estimates for a simulated ideal listener.
# Unlike for the analysis of participants' responses, there is no need to use a psychometric
# model since ideal observers have no attentional lapses (i.e., lambda = 0).
d.IO_intercept.slope.PSE <-
    # Assess the ideal observers at the exact same locations used during test blocks
    get_logistic_parameters_from_model(
      model = io,
      model_col = "io",
      groups = "Condition.Exposure") %>%
      crossing(Block = c(1:9)) %>%
  mutate(Phase = ifelse(Block %in% c(2, 4, 6), "exposure", "test")) %>%
  select(Condition.Exposure, Phase, Block, intercept_unscaled, slope_unscaled, intercept_scaled, slope_scaled, PSE) %>%
  add_block_labels() %>%
  mutate(across(c(Phase, Condition.Exposure, Block, Block.plot_label), factor))
```

```{r prepare-plot-intercepts-slopes}
p.across_blocks <-
  d.psychometric_intercept.slope.PSE %>%
  ggplot(
    aes(
      x = Block.plot_label, y = Intercept,
      ymin = Intercept.lower, ymax = Intercept.upper,
      colour = Condition.Exposure, group = Condition.Exposure)) +
  geom_rect(
    data = tibble(
      xmin = c(seq(0.5, 6.5, 2), seq(7.6, 8.6, 1)),
      xmax = c(seq(1.5, 7.5, 2), seq(8.5, 9.5, 1))),
    mapping = aes(xmin = xmin, xmax = xmax),
    ymin = -Inf, ymax = Inf, fill = "grey", alpha = .1, inherit.aes = F) +
  geom_point(position = position_dodge(.3), size = 1) +
  geom_linerange(linewidth = .6, position = position_dodge(.3), alpha = .5) +
  stat_summary(geom = "line", position = position_dodge(.3)) +
  scale_x_discrete("Block") +
  scale_colour_manual("Condition",
    labels = c("baseline", "+10ms", "+40ms", "prior"),
    values = c(colours.condition, "darkgray"),
    aesthetics = "color") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 22.5, hjust = 1))

p.intercept_1to7 <-
  p.across_blocks +  
  geom_hline(
    data = d.IO_intercept.slope.PSE,
    # Using scaled intercept (which is keeping the location at which the intercept is evaluated constant
    # across exposure and test: it's always the mean VOT of test)
    aes(yintercept = intercept_scaled, colour = Condition.Exposure, group = Condition.Exposure),
    linetype = 2,
    linewidth = .7,
    alpha = 0.3,
    inherit.aes = F) +
  scale_y_continuous("Intercept")

# get the relative distance of the IO-predicted PSEs from the starting point PSE at test 1
d.true_shift <-
  d.psychometric_intercept.slope.PSE %>%
  # Join the PSEs from ideal observers
  left_join(
    d.IO_intercept.slope.PSE %>%
      dplyr::select(Phase, Condition.Exposure, Block, PSE) %>%
      mutate(PSE = round(PSE, 1)) %>%
      rename(PSE.io_predicted = PSE), by = c("Condition.Exposure", "Block")) %>%
  group_by(Condition.Exposure) %>%
  mutate(
    PSE_block1 =
      case_when(
        Condition.Exposure == "Shift0" ~ d.psychometric_intercept.slope.PSE$PSE[1],
        Condition.Exposure == "Shift10" ~ d.psychometric_intercept.slope.PSE$PSE[10],
        Condition.Exposure == "Shift40" ~ d.psychometric_intercept.slope.PSE$PSE[19]),
    true_shift = PSE.io_predicted - PSE_block1,
    proportion_shift = (PSE - PSE_block1)/true_shift) %>%
  add_block_labels()

p.PSE_1to7 <-     
  p.across_blocks +  
  scale_y_continuous("PSE (ms VOT)") +
  aes(y = PSE, ymin = PSE.lower, ymax = PSE.upper) +
  geom_rect(
    data = d.IO_intercept.slope.PSE %>%
      group_by(Condition.Exposure) %>%
      summarise(PSE = mean(PSE)) %>%
      filter(!Condition.Exposure %in% c("Shift0", "Shift10", "Shift40", "prior")),
    mapping = aes(xmin = -Inf, ymin = min(PSE), ymax = max(PSE), xmax = Inf),
    fill = "#d2d4dc",
    alpha = .08,
    inherit.aes = F) +
  geom_linerange(linewidth = .6, position = position_dodge(.3), alpha = .5) +
  geom_hline(
    data = d.IO_intercept.slope.PSE %>% filter(Condition.Exposure %in% c("Shift0", "Shift10", "Shift40")),
    mapping = aes(yintercept = PSE, color = Condition.Exposure, group = Condition.Exposure),
    linetype = 2,
    linewidth = .8,
    alpha = 0.4) +
  geom_text(
    data = d.true_shift %>%
      filter(Phase == "exposure") %>%
      group_by(Condition.Exposure, true_shift) %>%
      summarise(),
    mapping = aes(
      x = 10.2, y = c(25, 35, 65),
      label = paste(round(true_shift), "(100%)"),
      colour = Condition.Exposure),
    colour = colours.condition,
    fontface = "bold",
    size = 3,
    inherit.aes = F) +
   coord_cartesian(xlim = c(0.75, 9), clip = "off") +
   theme(plot.margin = unit(c(1, 3.2, 1, 1), "lines"), legend.position = "none")
```



```{r plot-test-exposure-fit, fig.width=11, fig.height=3, message=FALSE}
cond_fit_test_exposure <-
  tibble(
    full_join(
      get_conditional_effects(fit_exposure, d.for_analysis, "exposure") %>%
        .[[1]] %>%
        mutate(Item.VOT = descale(VOT_gs, VOT.mean_test, VOT.sd_exposure)),
      get_conditional_effects(fit_test, d.for_analysis, "test") %>%
        .[[1]] %>%
        mutate(Item.VOT = descale(VOT_gs, VOT.mean_test, VOT.sd_test)))) %>%
  arrange(as.numeric(as.character(Block))) %>%
  mutate(
    Phase = ifelse(Block %in% c(2, 4, 6), "exposure", "test")) %>%
  add_block_labels()

label_colour <-
  tibble(
    Block.plot_label = c("Test 1", "Exposure 1", "Test 2", "Exposure 2", "Test 3", "Exposure 3", "Test 4"),
    Block.colour = c("grey", "white", "grey", "white", "grey", "white", "grey")) %>%
  mutate(Block.plot_label = factor(Block.plot_label, levels = Block.plot_label, ordered = T))

p.fit_1to7 <-
  cond_fit_test_exposure %>%
  filter(Block %in% c(1:7))  %>%
  ggplot() +
  geom_rect(
    data = label_colour,
    aes(
      xmin = -Inf, xmax = Inf,
      ymin = 1.05, ymax = 1.3,
      fill = Block.colour), show.legend = F) +
  scale_fill_manual(values = c("grey" = "grey", "white" = "white")) +
  new_scale_fill() +
  geom_linefit(
    data = d.for_analysis %>%
      group_by(Phase, Condition.Exposure, Block.plot_label) %>%
      filter(Block %in% c(1:7), Item.Labeled == FALSE),
    x = Item.VOT,
    y = estimate__,
    fill = NA,
    legend.position = "top",
    legend.justification = "right") +
  coord_cartesian(clip = "off", ylim = c(0, 1))

p.fit_7to9 <-
  cond_fit_test_exposure %>%
  filter(Block %in% c(7:9)) %>%
  ggplot() +
  geom_linefit(
    data = d.for_analysis %>%
      group_by(Condition.Exposure, Block) %>%
      filter(Block %in% c(7:9)),
    x = Item.VOT,
    y = estimate__,
    fill = "grey",
    legend.position = "none") +
  theme(axis.title.y = element_blank())
```

## Conceptual replication (averaging over test blocks)
<!-- TO DO: consider a much simpler result plot here that just shows the marginal effects across blocks (a) in terms of the categorization function and (b) in terms of the PSE. Then we could move the main figure and its description into the overview of incremental analyses, closer to where they are actually being discussed. -->
We first use to the psychometric mixed-effects model to analyze participants' behavior when averaging over all test blocks. This analysis recasts within a psychometric model the type of analysis that remains the most common in the field: it assesses overall changes in listeners' behavior after exposure, without telling us how these changes accumulate, how they relate to listeners' prior expectations, or how they compare to behavior that would be expected from a learner that has fully adapted to the unfamiliar talker. Prediction (2b) states that changes in listeners' categorization function should depend on the distribution of phonetic cues in the exposure input. Specifically, the +10 condition should elicit a rightward shift in the categorization function relative to the baseline condition, and the +40 condition should elicit an even larger rightward shift. This is also what previous work found [@kleinschmidt-jaeger2016].

Across all test blocks, participants were more likely to respond "t" the longer the VOT (`r get_bf(fit_test, "mu2_VOT_gs > 0")`). Exposure affected participants' categorization responses in the predicted direction. Marginalizing over Tests 1-6, participants in the +40 condition were less likely to respond "t" than participants in the +10 condition (`r get_bf(fit_test, "mu2_Condition.Exposure_Shift40vs.Shift10 < 0")`) or the baseline condition (`r get_bf(fit_test, "mu2_Condition.Exposure_Shift40vs.Shift10 + mu2_Condition.Exposure_Shift10vs.Shift0 < 0")`). There was also evidence---albeit less decisive---that participants in the +10 condition were less likely to respond "t" than participants in the baseline condition (`r get_bf(fit_test, "mu2_Condition.Exposure_Shift10vs.Shift0 < 0")`). That is, the +10 and +40 conditions resulted in categorization functions that were shifted rightwards compared to the baseline condition, as also evident in Figure \@ref(fig:plot-fit-PSE)A.^[The perceptual model contained in the our psychometric mixed-effects model describes the effect of VOT on the log-odds of "t"-responses as a line. The main effect of VOT is the average slope of that line across exposure conditions. The $\hat{\beta}$s for the comparisons across conditions indicate differences in the intercept of that line. Negative $\hat{\beta}$s thus indicate a *down*ward shift of that line in one condition, relative to the other. These downward shifts result in *right*ward shifts of the point of subjective equality (PSE), the VOT at which "t" and "d" responses are equally likely. This also shows in Figure \@ref(fig:plot-fit-PSE)A. In this figure, predictions are transformed into proportion "t"-responses and the downward shifts appear visually as a rightward shifts of the S-shaped categorization function (of one condition relative to another).]

Unlike the differences in the relative shift of the categorization function, there was little evidence that the *slopes* of listeners' categorization functions differed between exposure conditions (0.7 < BFs < 2.2; see also Figure \@ref(fig:plot-fit-PSE)A). This lack of notable differences in the slope is precisely what is expected under distributional learning models, since our exposure conditions manipulated neither the category variances of /d/ and /t/ nor the distance between their category means. In the remainder of the main text, we thus focus on 'shifts' in the categorization function---i.e., changes in listeners' PSE. Parallel analyses of changes (or lack thereof) in the slopes of listeners' categorization functions are reported in the SI (\@ref(sec:slopes-analyses-test)-\@ref(sec:slopes-analyses-exposure)), and do not affect any of our conclusions.

In summary, the analysis across all test blocks conceptually replicates previous findings that exposure to different VOT distributions changes listeners' categorization responses [for /b/-/p/: @clayards2008; @kleinschmidt2015; @kleinschmidt2020; for /g/-/k/, @theodore-monto2019]. This replication is obtained for stimuli with phonetic properties that more closely resemble those experienced in everyday speech perception. This conceptual replication also adds to a small body of work that goes beyond dichotomous comparisons of exposure conditions, testing stronger hypotheses about the *relative order* of exposure effects [e.g., @babel2019; @bejjanki2011; @bradlow-bent2008; @cummings-theodore2023; @kleinschmidt2020; @liu-jaeger2018].

## Overview of incremental analyses
Next, we turn to the questions of primary interest. Incremental changes in participants' categorization responses can be assessed from three mutually complementing perspectives. First, we compare how exposure affects listeners' categorization responses *relative to other exposure conditions*. This is the perspective taken in previous studies and in the conceptual replication presented above, but extended to test *how early* in the experiment differences between exposure conditions begin to emerge. Second, we compare how exposure *incrementally changes* listeners' categorization responses from block to block within each condition, relative to listeners' responses prior to any exposure. Together with the first perspective, this allows us to test predictions (1)-(2a,b) from the introduction. Third, we compare changes in listeners' responses to those expected from an ideal observer that has fully learned the exposure distributions. This analysis has the potential to identify constraints on cumulative adaptation. As we show, in particular these latter two perspectives---made possible by the incremental exposure-test paradigm---afford stronger tests of predictions (3 - *learn to convergence*) and (4 - *diminishing returns*), and suggest previously unrecognized constraints on the early moments of adaptive speech perception. For all three analyses, we initially focus on Tests 1-4 with intermittent exposure. Following that, we analyze the effects of repeated testing without intermittent exposure blocks during Tests 4-6. Though research---including some of our own previous work---tends to interpret tests as passive windows into the effects of exposure, test stimuli constitute part of the exposure input listeners' receive. This has both methodological and theoretical consequences: for example, we show below that analyses like those used in previous work (and in our conceptual replication) tend to underestimate the true effect of exposure.

(ref:plot-fit-PSE) Summary of results. **Panel A:** Changes in listeners psychometric categorization functions as a function of exposure, from Test 1 to Test 4 with all intervening exposure blocks (only unlabeled trials were included in the analysis of exposure blocks since labeled trials provide little information about listeners' categorization function). Point ranges indicate the mean proportion of participants' "t"-responses and their 95% bootstrapped CIs. Lines and shaded intervals show the *maximum a posteriori* (MAP) estimates and 95%-CIs of a Bayesian mixed-effects psychometric model fit to participants' responses. **Panel B:** Same as Panel A but for the final three test blocks without intervening exposure. Test 4 is shown as part of both Panels A and B. **Panel C:** Changes across blocks and conditions in listeners' point-of-subjective-equality (PSE) of the lapse-corrected categorization functions from Panels A \& B (i.e., the PSE of the perceptual model inferred from listeners' responses; for changes in the slope of that function, see SI, \@ref(sec:io-bias-correction)). Point ranges represent the posterior medians and their 95%-CIs derived from the psychometric model. Horizontal dashed lines indicate 95%-CIs of the PSEs expected from an idealized learner (an ideal observer model that has fully learned the exposure distributions). Percentage labels indicate the amount of shift in PSE exhibited by participants as a proportion of the expected shift under the idealized learners. Horizontal gray ribbon indicates the 95%-CIs of the PSEs expected from an idealized listener *prior to any exposure*.

\begin{landscape}

```{r plot-fit-PSE, fig.width=12.5, fig.height=7, fig.cap="(ref:plot-fit-PSE)"}
p.fit_1to7 + 
  p.fit_7to9 + 
  (p.PSE_1to7 +
     geom_label(
       data = d.true_shift %>%
         filter(Block != 1),
       mapping = aes(label = paste(scales::percent(proportion_shift, accuracy = .1))),
       position = position_dodge(.3),
       label.size = .15,
       label.padding = unit(.18, "lines"),
       size = 2.5,
       show.legend = F)) +
  plot_layout(
    design = "
AAAAAAAAA
BBBDDDDDD
###DDDDDD
",
heights = c(1.1, 1.1, 1.8)) +
  plot_annotation(tag_levels = 'A', tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))
```

\end{landscape}

```{r}
# Not removing the PSE panel since we replot and augment it further down
rm(p.across_blocks, p.estimates_1to7, p.fit_1to7, p.fit_7to9, label_colour, 
   fit_test_nested_within_condition_and_block, fit_exposure, fit_exposure_nested, cond_fit_test_exposure)
```

## How quickly does exposure affect listeners' categorization responses? (comparing exposure conditions within each test block)

```{r fit-simple-effects-condition, results='hide'}
# Get simple effects of Condition nested under block
my_priors <- c(
  prior(student_t(3, 0, 2.5), class = "b", dpar = "mu2"),  
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block1:VOT_gs"),
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block3:VOT_gs"),
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block5:VOT_gs"),
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block7:VOT_gs"),
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block8:VOT_gs"),
  set_prior("student_t(3, 0, 15)", dpar = "mu2", coef = "Block9:VOT_gs"),
  prior(cauchy(0, 2.5), class = "sd", dpar = "mu2"),
  prior(lkj(1), class = "cor"))

fit_test.simple_effects_condition <-
  brm(
  bf(
    Response.Voiceless ~ 1,
    mu1 ~ 0 + offset(0),
    mu2 ~ 0 + Block / (Condition.Exposure * VOT_gs) +
      (0 + Block / VOT_gs | ParticipantID) +
      (0 + Block / (Condition.Exposure * VOT_gs) | Item.MinimalPair),
    theta1 ~ 1),
    data = d.for_analysis %>%
      filter(Phase == "test") %>%
      prepVars(levels.Condition = levels_Condition.Exposure, contrast_type = contrast_type),
    priors = my_priors,
    cores = 4,
    chains = chains,
    init = 0,
    iter = 4000,
    warmup = 2000,
    family = mixture(bernoulli("logit"), bernoulli("logit"), order = F),
    sample_prior = "yes",
    control = list(adapt_delta = .995),
    file ="../models/test-nested-condition.rds")
```

```{r hypothesis-table-simple-effects-condition, results='asis'}
hyp.simple_effects_condition <-
  hypothesis(
    fit_test.simple_effects_condition,
    c(
      "mu2_Block1:Condition.Exposure_Shift10vs.Shift0 = 0",
      "mu2_Block1:Condition.Exposure_Shift40vs.Shift10 = 0",
      "mu2_Block1:Condition.Exposure_Shift40vs.Shift10 + mu2_Block1:Condition.Exposure_Shift10vs.Shift0 = 0",
      "mu2_Block3:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block3:Condition.Exposure_Shift40vs.Shift10 < 0",
      "mu2_Block3:Condition.Exposure_Shift40vs.Shift10 + mu2_Block3:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block5:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block5:Condition.Exposure_Shift40vs.Shift10 < 0",
      "mu2_Block5:Condition.Exposure_Shift40vs.Shift10 + mu2_Block5:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block7:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block7:Condition.Exposure_Shift40vs.Shift10 < 0",
      "mu2_Block7:Condition.Exposure_Shift40vs.Shift10 + mu2_Block7:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block8:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block8:Condition.Exposure_Shift40vs.Shift10 < 0",
      "mu2_Block8:Condition.Exposure_Shift40vs.Shift10 + mu2_Block7:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block9:Condition.Exposure_Shift10vs.Shift0 < 0",
      "mu2_Block9:Condition.Exposure_Shift40vs.Shift10 < 0",
      "mu2_Block9:Condition.Exposure_Shift40vs.Shift10 + mu2_Block7:Condition.Exposure_Shift10vs.Shift0 < 0"),
    robust = T) %>%
  .$hypothesis %>%
  dplyr::select(-Star)

make_hyp_table(
  fit_test.simple_effects_condition,
  hyp.simple_effects_condition,
  c("$PSE_{+10} = PSE_{baseline}$", 
    "$PSE_{+40} = PSE_{+10}$", 
    "$PSE_{+40} = PSE_{baseline}$", 
    rep(c("$PSE_{+10} > PSE_{baseline}$", 
          "$PSE_{+40}  > PSE_{+10}$", 
          "$PSE_{+40} > PSE_{baseline}$"), 5)),
    caption = "When did exposure begin to affect participants' categorization responses? When, if ever, were these changes undone with repeated testing? This table summarizes the simple effects of the exposure conditions for each test block. Note that rightward shifts of the categorization function (and its PSE) correspond to negative estimates (lower intercepts in predicting the log-odds of \`\`t\'\'-responses). Null predictions (e.g., for Test 1) were tested using the Savage-Dickey density ratio [@wagenmakers2010].") %>%
  pack_rows("Test block 1 (pre-exposure)", 1, 3) %>%
  pack_rows("Test block 2", 4, 6) %>%
  pack_rows("Test block 3", 7, 9) %>%
  pack_rows("Test block 4", 10, 12) %>%
  pack_rows("Test block 5 (repeated testing without additional exposure)", 13, 15) %>%
  pack_rows("Test block 6 (repeated testing without additional exposure)", 16, 18)

rm(fit_test.simple_effects_condition)
```

Figure \@ref(fig:plot-fit-PSE)A suggests that differences between exposure conditions emerged early in the experiment: already in Test 2, listeners in the +10 condition have shifted their categorization functions rightwards relative to the baseline condition, and listeners in the +40 condition have shifted their in categorization functions even further rightwards. This is confirmed by Bayesian hypothesis tests summarized in Table \@ref(tab:hypothesis-table-simple-effects-condition), which we go through next. 

Prior to any exposure, during Test 1, participants' responses did not differ across exposure condition. This result is predicted by models of adaptive speech perception under the assumptions that (a) participants in the different groups have similar prior experiences, and that (b) our sample size is sufficiently large to yield stable estimates of listeners' categorization function.^[<!-- This comparatively simple procedure estimates the BF as the ratio of the posterior and prior density over the null. --> The moderate BFs for these hypothesis tests are due to our use of regularizing priors, which have non-negligible density over the null. Rather than to test the null against more plausible alternative priors, we appeal to readers' intuition: the 90%-CIs of the comparisons for Test 1 are all wide and approximately centered around zero. <!-- This means there is very little evidence in favor of an effect in either direction. -->] Equality of pre-exposure behavior across exposure groups is also implicitly assumed---but rarely tested---in the interpretation of most studies on adaptive speech perception [when it is tested, it often turns out that this assumption is *not* necessarily warranted, presumably due to insufficient sample sizes, cf. @kleinschmidt2020]. 

During Test 2, after exposure to only 24 /d/ and 24 /t/ stimuli (thereof half labeled), participants' categorization responses already differed between exposure conditions (BFs > 14). All differences between exposure conditions that emerged at Test 2 followed prediction (2b - *exposure distributions*). Additional analyses reported in the SI (\@ref(sec:SI-exposure-block-analysis)) found that listeners' categorization functions had already changed in the predicted direction during the first *exposure* block, in line with Figure \@ref(fig:plot-fit-PSE)A. This suggests that changes in listeners' categorization responses emerged quickly at the earliest point tested---after only a fraction of exposure trials previously tested in similar paradigms.

The effects of the three exposure conditions continued to persist until Test 4, always in line with prediction (2b). Table \@ref(tab:hypothesis-table-simple-effects-condition) does, however, indicate an interesting non-monotonic development in the way that listeners' categorization function changed. While the difference between the +40 condition and both the baseline and +10 condition continued to increase numerically with increasing exposure (increasingly larger magnitude of negative estimates in Tests 2-4), the same was not the case for the difference between the +10 and the baseline condition. Instead, the difference between the +10 and baseline condition *reduced* with increasing exposure (while maintaining its direction; from -1.45 to -0.88, see Table \@ref(tab:hypothesis-table-simple-effects-condition)). <!-- TO DO: replace -1.45 and -0.88 with R expression using get_bf function, modifying it to just get the estimate (or some other reasonable approach) --> At first blush, this non-monotonicity would potentially constitute evidence against prediction (2a) that the magnitude of exposure effects should *in*crease with increasing exposure. In the next section, we show that this would be the wrong conclusion to draw because it fails to take into consideration how listeners' behavior changes *relative to their prior expectations*. Indeed, the seemingly unexpected non-monotonicity---which would be impossible to detect without repeated testing---turns out to be important in understanding incremental adaptation.

## Incremental adaptation from prior expectations (comparing block-to-block changes within exposure conditions)
Next, we compare how exposure affected listeners' categorization responses from block to block *within* each exposure condition. This allows us to understand changes in listeners' categorization function relative to listeners' pre-exposure behavior, thereby assessing the joint effects of predictions (1 - *prior expectations*) and (2a,b - *exposure amount & distributions*). To facilitate visual comparison across blocks and conditions, Figure \@ref(fig:plot-fit-PSE)C summarizes the block-to-block changes in listeners' PSE. Focusing for now on Tests 1-4, this highlights three aspects of participants' behavior that were not readily apparent in the statistical comparisons presented so far.

```{r fit-simple-effects-block, results='hide'}
my_priors <- c(
  prior(student_t(3, 0, 2.5), class = "b", dpar = "mu2"),
  prior(student_t(3, 0, 15), class = "b", dpar = "mu2", coef = "Condition.ExposureShift0:VOT_gs"),
  prior(student_t(3, 0, 15), class = "b", dpar = "mu2", coef = "Condition.ExposureShift10:VOT_gs"),
  prior(student_t(3, 0, 15), class = "b", dpar = "mu2", coef = "Condition.ExposureShift40:VOT_gs"),
  prior(cauchy(0, 2.5), class = "sd", dpar = "mu2"),
  prior(lkj(1), class = "cor"))

fit_test.simple_effects_block <-
  brm(
  bf(
    Response.Voiceless ~ 1,
    mu1 ~ 0 + offset(0),
    mu2 ~ 0 + Condition.Exposure/(Block * VOT_gs) +
      (0 + Block * VOT_gs | ParticipantID) +
      (0 + Condition.Exposure/(Block * VOT_gs) | Item.MinimalPair),
    theta1 ~ 1),
    data = d.for_analysis %>%
      filter(Phase == "test") %>%
      prepVars(levels.Condition = levels_Condition.Exposure, contrast_type = contrast_type),
    priors = my_priors,
    cores = 4,
    chains = chains,
    init = 0,
    iter = 4000,
    warmup = 2000,
    family = mixture(bernoulli("logit"), bernoulli("logit"), order = F),
    sample_prior = "yes",
    control = list(adapt_delta = .995),
    file ="../models/test-nested-block.rds")
```

```{r hypothesis-table-simple-effects-block, results='asis'}
hyp.simple_effects_block <-
  hypothesis(
    fit_test.simple_effects_block,
    c(
      "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 > 0",
      "mu2_Condition.ExposureShift0:Block_Test3vs.Test2 > 0",
      "mu2_Condition.ExposureShift0:Block_Test4vs.Test3 > 0",
      "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 + mu2_Condition.ExposureShift0:Block_Test3vs.Test2 + mu2_Condition.ExposureShift0:Block_Test4vs.Test3 > 0",
      "mu2_Condition.ExposureShift0:Block_Test5vs.Test4 < 0",
      "mu2_Condition.ExposureShift0:Block_Test6vs.Test5 < 0",                     
      "mu2_Condition.ExposureShift0:Block_Test5vs.Test4 + mu2_Condition.ExposureShift0:Block_Test6vs.Test5 < 0",

      "mu2_Condition.ExposureShift10:Block_Test2vs.Test1 > 0",
      "mu2_Condition.ExposureShift10:Block_Test3vs.Test2 > 0",
      "mu2_Condition.ExposureShift10:Block_Test4vs.Test3 > 0",
      "mu2_Condition.ExposureShift10:Block_Test2vs.Test1 + mu2_Condition.ExposureShift10:Block_Test3vs.Test2 + mu2_Condition.ExposureShift10:Block_Test4vs.Test3 > 0",

      "mu2_Condition.ExposureShift10:Block_Test5vs.Test4 < 0",
      "mu2_Condition.ExposureShift10:Block_Test6vs.Test5 < 0",
      "mu2_Condition.ExposureShift10:Block_Test5vs.Test4 + mu2_Condition.ExposureShift10:Block_Test6vs.Test5 < 0",

      "mu2_Condition.ExposureShift40:Block_Test2vs.Test1 < 0",
      "mu2_Condition.ExposureShift40:Block_Test3vs.Test2 < 0",
      "mu2_Condition.ExposureShift40:Block_Test4vs.Test3 < 0",
      "mu2_Condition.ExposureShift40:Block_Test2vs.Test1 + mu2_Condition.ExposureShift40:Block_Test3vs.Test2 + mu2_Condition.ExposureShift40:Block_Test4vs.Test3 < 0",

      "mu2_Condition.ExposureShift40:Block_Test5vs.Test4 > 0",
      "mu2_Condition.ExposureShift40:Block_Test6vs.Test5 > 0",
      "mu2_Condition.ExposureShift40:Block_Test5vs.Test4 + mu2_Condition.ExposureShift40:Block_Test6vs.Test5 > 0"),
    robust = T) %>%
  .$hypothesis %>%
  dplyr::select(-Star)

make_hyp_table(
  fit_test.simple_effects_block,
  hyp.simple_effects_block,
    c(rep(c(
  # Comparing differences blocks within conditions
  "Block 1 to 2: decreased PSE",
  "Block 2 to 3: decreased PSE",
  "Block 3 to 4: decreased PSE",
  "{\\em Block 1 to 4: decreased PSE}",
  "Block 4 to 5: increased PSE",
  "Block 5 to 6: increased PSE",
  "{\\em Block 4 to 6: increased PSE}"), 2),
  "Block 1 to 2: increased PSE",
  "Block 2 to 3: increased PSE",
  "Block 3 to 4: increased PSE",
  "{\\em Block 1 to 4: increased PSE}",
  "Block 4 to 5: decreased PSE",
  "Block 5 to 6: decreased PSE",
  "{\\em Block 4 to 6: decreased PSE}"),
    caption = "Was there incremental change from test block 1 to 4? Did these changes dissipate with repeated testing from block 4 to 6? This table summarizes the simple effects of block for each exposure condition. Note that rightward shifts of the categorization function (and its PSE) correspond to negative estimates (lower intercepts in predicting the log-odds of \`\`t\'\'-responses).") %>%
  pack_rows("Difference between blocks: baseline", 1, 7) %>%
  pack_rows("Difference between blocks: +10", 8, 14) %>%
  pack_rows("Difference between blocks: +40", 15, 21)
```

First, while the PSEs for the +40 and +10 conditions were shifted rightwards compared to the baseline condition, both the +10 and the baseline condition seem to shift *left*wards relative to their pre-exposure starting point in Test 1. Bayesian hypothesis tests summarized in Table \@ref(tab:hypothesis-table-simple-effects-block) find moderate support for a leftward shift from Test 1 to 4 in both the +10 condition (BF = 3.5) and the baseline condition (BF = 7.6). In contrast, there was strong support that the +40 condition shifted rightwards relative to pre-exposure (BF = 45.2). <!-- TO DO: can we use a modified get_bf function to report these BFs, rather than to hand-code them? --> To understand this pattern, it is helpful to relate the three exposure conditions to the phonetic distribution in listeners' prior experience. Figure \@ref(fig:exposure-means-database-matrix-plot) shows the exposure means for /d/ and /t/ relative to the distributions of three important cues to the word-initial /d/-/t/ contrast in L1-US English [based on databases of isolated and connected speech, @chodroff-wilson2018]. This comparison offers an explanation as to why the baseline condition (and to some extent the +10 condition) shift leftwards with increasing exposure, whereas the +40 condition shifts rightwards: relative to the distribution of VOT for /d/ and /t/ in listeners' prior experience, only the +40 condition presents category means that are clearly larger than expected along VOT, whereas the baseline condition and, to some extent, the +10 condition presented lower-than-expected category means. That is, once we take into account how our exposure conditions relate to listeners' prior experience (prediction 1), both the direction of changes from Test 1 to 4 *within* each exposure condition (Table \@ref(tab:hypothesis-table-simple-effects-block)), and the direction of differences *between* exposure conditions receive an explanation (Table \@ref(tab:hypothesis-table-simple-effects-condition)). To further illustrate this point, the horizontal gray ribbon in  \@ref(fig:plot-fit-PSE)C shows the range of PSEs predicted by Bayesian ideal observers trained on the distribution of VOT, f0, and vowel duration for isolated word productions in Figure \@ref(fig:exposure-means-database-matrix-plot) (for details, see SI, \@ref(sec:idealized-prior-listeners)).

(ref:exposure-means-database-matrix-plot) Placement of exposure stimuli relative to an estimate of typical phonetic distributions for `r nrow(d.chodroff_wilson)` word-initial /d/ and /t/ productions by `r length(unique(d.chodroff_wilson$Talker))` female L1 talkers of US English in @chodroff-wilson2018. After voice onset time, f0 and vowel duration are two of the most informative cues to word-initial /d/-/t/ in L1 US English. For details, see SI \@ref(sec:idealized-prior-listeners). Colored labels show the category means of the exposure conditions. 

```{r exposure-means-database-matrix-plot, fig.width=base.width*4, fig.height=base.height*2.5+1, fig.cap="(ref:exposure-means-database-matrix-plot)", fig.pos="!h"}
d.chodroff_wilson %>%
  group_by(speechstyle, category) %>%
  filter(if_all(c(VOT, f0_Mel, vowel_duration), ~ abs(scale(.x)[,1]) < 3.5)) %>%
  ggplot(aes(color = category, fill = category)) +
  geom_autopoint(aes(shape = speechstyle), alpha = .4, show.legend = F) +
  geom_autodensity(aes(linetype = speechstyle), position = "identity", alpha = .5) +
  stat_ellipse(aes(x = .panel_x, y = .panel_y, linetype = speechstyle)) +
  scale_color_manual(
    "Category",
    labels = c("/d/", "/t/"),
    values = colours.category_greyscale,
    aesthetics = c("color", "fill")) +
  new_scale_color() +
  geom_text(
    data =
      d.for_analysis %>%
      filter(Phase == "exposure") %>%
      group_by(Condition.Exposure, category) %>%
      summarise(across(c(VOT, f0_Mel, vowel_duration), mean)),
    mapping = aes(x = .panel_x, y = .panel_y, label = category, colour = Condition.Exposure),
    alpha = .8, size = 3,
    show.legend = F,
    inherit.aes = F) +
  geom_rug(
    d.for_analysis %>%
      filter(Phase == "exposure") %>%
      group_by(Condition.Exposure, category) %>%
      summarise(across(c(VOT, f0_Mel, vowel_duration), mean)),
    mapping = aes(x = .panel_x, colour = Condition.Exposure),
    sides = "b") +
  guides(
    colour = "none",
    linetype = guide_legend(title = "Speech style", override.aes = list(fill = NA, shape = c(0, 3)))) +
  scale_colour_manual(
    labels = c("baseline", "+10ms", "+40ms"),
    values = colours.condition) +
  facet_matrix(
    vars(c(VOT, f0_Mel, vowel_duration)),
    layer.lower = c(3:5), layer.diag = c(2,6), layer.upper = c(1, 4:5),
    labeller = labeller(
      .rows = c(VOT = "VOT\n(ms)", f0_Mel = "f0\n(Mel)", vowel_duration = "Vowel duration\n(ms)"),
      .cols = c(VOT = "VOT\n(ms)", f0_Mel = "f0\n(Mel)", vowel_duration = "Vowel duration\n(ms)"))) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r prediction-error-during-exposure}
d.exposure_surprisal <-
  d.for_analysis %>%
  filter(Phase == "exposure", Item.Labeled) %>%
  cross_join(
    io %>%
      filter(Condition.Exposure %in% c("Shift0", "Shift10", "Shift40", "prior")) %>%
      select(-x, fold) %>%
      rename(IO.Type = Condition.Exposure)) %>%
  # For each labeled exposure stimulus, calculate the surprisal of seeing the category label
  # (this is not quite parallel to what we did for test tokens, which was looking at the
  # surprisal of the VOT, rather than the surprisal of the category label)
  mutate(
    x = ifelse(IO.Type != "prior", Item.VOT, pmap(list(Item.VOT, Item.f0_Mel, Item.VowelDuration), ~ c(...))),
    posterior = pmap_dbl(
      .l =
        list(
          x,
          category,
          io),
      .f = function(x, y, z)
        pmap(
          list(z$mu, z$Sigma, z$Sigma_noise),
          ~ dmvnorm(x, ..1, ..2 + ..3)) %>%
        reduce(~ ..1 / (..1 + ..2)) %>%
        { if (y == "/d/") . else 1 - .}),
    surprisal = -log2(posterior))

d.exposure_surprisal %<>%
  group_by(Condition.Exposure, IO.Type) %>%
  summarise(across(surprisal, list(mean = mean, sd = sd)))
```

Second, we find support for prediction (4) about the *diminishing returns* of additional exposure predicted by some theories of adaptive speech perception. The estimates in Table \@ref(tab:hypothesis-table-simple-effects-block) suggest that listeners' PSEs changed most from Test 1 to Test 2, and then changed less and less with additional exposure up to Test 4 (smaller magnitude of estimates compared to earlier test blocks). This seems to be particularly pronounced for the two conditions that exhibited the largest shifts relative to pre-exposure, the baseline condition and the +40 condition. As mentioned in our Open Science statement, our experiment was not designed to have high power to assess such *changes in the magnitude of the shifts* across the block within each condition. We did, however, conduct post-hoc hypothesis tests to assess the support for this pattern. These tests found anecdotal to moderately strong evidence in support of prediction (4 - *diminishing returns*). For the +40 condition, the shift from Test 1 to 2 was larger than the shift from Test 2 to 3 (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift40:Block_Test2vs.Test1 < mu2_Condition.ExposureShift40:Block_Test3vs.Test2")`), which was larger than the shift from Test 3 to 4 (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift40:Block_Test3vs.Test2 < mu2_Condition.ExposureShift40:Block_Test4vs.Test3")`). Comparing the change from Test 1 to 2 against the change from Test 3 to 4, there was stronger support that the speed of changes in the PSE decreased (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift40:Block_Test2vs.Test1 < mu2_Condition.ExposureShift40:Block_Test4vs.Test3")`). For the baseline condition, the shift from Test 1 to 2 was larger than the shift from Test 2 to 3 (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 > mu2_Condition.ExposureShift0:Block_Test3vs.Test2")`), which was almost identical, but slightly smaller, than the shift from Test 3 to 4 (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift0:Block_Test4vs.Test3 > mu2_Condition.ExposureShift0:Block_Test3vs.Test2")`). Again, a comparison of the change from Test 1 to 2 against the change from Test 3 to 4, yielded the strongest support that the speed of changes in the PSE decreased (`r get_bf(fit_test.simple_effects_block, "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 > mu2_Condition.ExposureShift0:Block_Test4vs.Test3")`). For both the +40 and the baseline condition, there was only anecdotal evidence that the final exposure block resulted in *any additional* shift in listeners' PSE (BFs $\leq$ 1.7, cf. Table \@ref(tab:hypothesis-table-simple-effects-block)). 

```{r, include=F}
# Assessing evidence for changes in the PSEs within each condition
hypothesis(
  fit_test.simple_effects_block,
  c(
    "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 > mu2_Condition.ExposureShift10:Block_Test3vs.Test2",
    "mu2_Condition.ExposureShift0:Block_Test2vs.Test1 + mu2_Condition.ExposureShift0:Block_Test3vs.Test2 + mu2_Condition.ExposureShift0:Block_Test4vs.Test3 > mu2_Condition.ExposureShift10:Block_Test2vs.Test1 + mu2_Condition.ExposureShift10:Block_Test3vs.Test2 + mu2_Condition.ExposureShift10:Block_Test4vs.Test3"),
  robust = T)
```

Third and finally, Panel C also begins to illuminate the reasons for the non-monotonic development of the +10 and baseline conditions relative to each other, discussed in the previous section. In particular, this non-monotonicity does *not* appear due to a reversal of the effects in either of the two exposure conditions. Rather, both exposure conditions continue to change listeners' categorization function in the same direction from Test 1 to Test 4, in line with predictions (2a) and (2b). However, after the rapid change from the pre-exposure Test 1 to the first post-exposure Test 2, listeners' categorization responses in the baseline condition did not change as much as in the +10 condition. In fact, listeners' categorization function in the baseline condition seems to have changed very little at all beyond the first exposure block. 

This explains the reduction in the difference between the +10 and baseline conditions discussed in the previous section. It does, however, raise the question *why* listeners' responses in the baseline condition did not change further with increasing exposure. One explanation would be that listeners in the baseline condition did for some reason---including chance---fully learn the relevant phonetic distributions within a single block of exposure, whereas listeners in the +10 condition were not. Our final perspective suggests that this was *not* the case.

## Constraints on cumulative adaptation (comparing exposure effects against idealized learner models)

```{r hypothesis-table-shrinkage, results='asis'}
fit_test_nested_within_condition_and_block <- readRDS("../models/test-nested_slope-priorSD15-0.995.rds")

predictedPSE_40 <- get_IO_predicted_PSE(condition = "Shift40")
predictedPSE_10 <- get_IO_predicted_PSE(condition = "Shift10")
predictedPSE_0 <- get_IO_predicted_PSE(condition = "Shift0")
predictedPSE_prior <- get_IO_predicted_PSE(condition = "prior")

hyp.shrinkage <-
  hypothesis(
    fit_test_nested_within_condition_and_block,
    c(
      str_c("abs(mu2_IpasteCondition.ExposureBlocksepEQxShift0x7 - ", predictedPSE_0, ") > 0" ),
      str_c("abs(mu2_IpasteCondition.ExposureBlocksepEQxShift10x7 - ", predictedPSE_10, ") > 0" ),
      str_c("abs(mu2_IpasteCondition.ExposureBlocksepEQxShift0x7 - ", predictedPSE_40, ") > 0" ),
      # TO DO: the next two lines need fixing. Same for the final two statements in the text below
      # See labels below for description of what SHOULD be done
      str_c("(mu2_IpasteCondition.ExposureBlocksepEQxShift10x7  -", predictedPSE_prior, ")/(", predictedPSE_10, "-", predictedPSE_prior, ") >", "(mu2_IpasteCondition.ExposureBlocksepEQxShift0x7  -", predictedPSE_prior, ")/(", predictedPSE_0, "-", predictedPSE_prior, ")"),
      str_c("(mu2_IpasteCondition.ExposureBlocksepEQxShift40x7  -", predictedPSE_prior, ")/(", predictedPSE_40, "-", predictedPSE_prior, ") >", "(mu2_IpasteCondition.ExposureBlocksepEQxShift0x7  -", predictedPSE_prior, ")/(", predictedPSE_0, "-", predictedPSE_prior, ")")),
    robust = T) %>%
  .$hypothesis %>%
  dplyr::select(-Star)

make_hyp_table(
  fit_test_nested_within_condition_and_block,
  hyp.shrinkage,
  c(
    "$|\\Delta(PSE_{ideal_{baseline}}, PSE_{actual_{baseline}})| > 0$",
    "$|\\Delta(PSE_{ideal_{+10}}, PSE_{actual_{+10}})| > 0$",
    "$|\\Delta(PSE_{ideal_{+40}}, PSE_{actual_{+40}})| > 0$",
    "$\\frac{\\Delta(PSE_{actual_{+10}}, PSE_{ideal_{pre}})}{\\Delta(PSE_{ideal_{+10}}, PSE_{ideal_{pre}})} > \\frac{\\Delta(PSE_{actual_{baseline}}, PSE_{ideal_{pre}})}{\\Delta(PSE_{ideal_{baseline}}, PSE_{ideal_{pre}})}$",
    "$\\frac{\\Delta(PSE_{actual_{+40}}, PSE_{ideal_{pre}})}{\\Delta(PSE_{ideal_{+40}}, PSE_{ideal_{pre}})} > \\frac{\\Delta(PSE_{actual_{baseline}}, PSE_{ideal_{pre}})}{\\Delta(PSE_{ideal_{baseline}}, PSE_{ideal_{pre}})}$"),
  caption = "Comparison of actual changes in participants' categorization function against those expected from idealized learners. This table summarizes results for the test block following the final exposure block (Test 4). For identical tests for all test blocks, see SI (\\ref{sec:XXX}).",
  col1_width = "25em") %>%
  pack_rows("Have participants converged against the PSE expected from idealized learner?", 1, 3) %>% 
  pack_rows("Does convergence against ideal PSE differ across conditions (asymmetric shrinkage)?", 4, 5)
```

Beyond the perspectives on incremental adaptation discussed so far, Figure \@ref(fig:plot-fit-PSE)C compares participants' responses against those of an idealized learner that has fully learned the exposure distributions (colored dashed lines). Specifically, we fit Bayesian ideal observers against the labeled VOT distributions of each exposure condition, using the same approach used for the idealized pre-exposure listeners (horizontal gray ribbons). The dashed lines show the PSEs expected from such idealized learners (for details, see SI \@ref(sec:io-bias-correction)). This approach follows previous work from our lab [@kleinschmidt-jaeger2016; @kleinschmidt2020], and makes it possible to assess how far listeners have converged against the exposure distributions. The relevant hypothesis tests are summarized in Table \@ref(tab:hypothesis-table-shrinkage).

Figure \@ref(fig:plot-fit-PSE)C suggests that listeners in all three exposure conditions did very likely *not* fully learn the exposure distributions ($|\Delta(PSE_{ideal}, PSE_{actual})| > 0$: all BFs $\ge 8000$). By itself, a failure to have converged against the performance of an idealized learner would not necessarily constitute evidence against prediction (3 - *learn to convergence*): listeners might simply not have received sufficient exposure to have learned the exposure distribution. However, as already described for the baseline condition, participants' behavior changed little at all after the first exposure block. Instead, participants seem to have prematurely converged against stable behavior long before they had fully learned the exposure distribution. 

The percentage labels in Figure \@ref(fig:plot-fit-PSE)C quantify the degree to which participants adapted their PSE towards the statistics of the exposure condition: 0% would correspond to no change relative to the listeners' PSE in Test 1, and 100% would correspond to the PSE predicted for an idealized learner who has fully converged against the exposure distributions. In the baseline condition, changes in participants' PSE seem to converge against approximately 20% of the PSE expected from an idealized learner. A similar pattern of premature convergence is evident for the +40 condition: changes in participants' PSE seem to have leveled off by Test 4, despite the fact that participants' PSE had shifted less than half way to the idealized learner's PSE. (For the +10 condition, it is less clear whether participants had already converged against a PSE.) That is, in terms of the possible adaptation scenarios depicted in Figure \@ref(fig:predictions) in the introduction, it seems that our results most closely resemble the scenario shown in the right-most column of panel B. Notably, this *premature convergence* has consequences for participants' recognition accuracy: while incremental adaptation substantially improved participants' recognition accuracy compared to their pre-exposure accuracy, only participants in the +10 condition came close to achieving the recognition accuracy expected from a learner that has fully converged against the exposure distributions. At least at first blush then, listeners in the baseline and +40 condition seem to have stopped adapting despite the fact that further adaptation would have further improved their recognition accuracy.^[While a failure to improve from, say, 90% and 95% accuracy might not seem noteworthy, it implies misunderstanding one in ten vs. one in twenty words, thus *doubling* the odds of recognition errors.]

(ref:IO-human-accuracy) Changes across blocks and conditions in participants' recognition accuracy for the unfamiliar talker's speech. For each block, we used participants' categorization functions---estimated by the psychometric mixed-effects model fit to participants' responses---to categorize all 144 exposure inputs of that exposure condition. Accuracy was calculated for the two decision rules that are most commonly assumed to underlie speech recognition as well as perceptual decision-making in other domains [for review, see @massaro-friedman1990]: Luce's choice rule (responding proportional to posterior probability of category) and the criterion choice rule (always responding with the category that has highest posterior probability). As in Figure \@ref(fig:plot-fit-PSE)C, point ranges represent the posterior medians and their 95%-CIs derived from the psychometric model. Horizontal dashed lines indicate accuracy expected from an idealized learner (an ideal observer model that has fully learned the exposure distributions) and horizontal shaded ribbons indicate the 95%-CI expected from an idealized pre-exposure listener. <!-- TO DO: provide ribbon based on CV rather than a single line -->

```{r compute-IO-human-accuracy}
# get io accuracy on shift conditions they were trained on
# prior io gets no training so prior-io-accuracy will be evaluated on all three conditions
io.accuracy <- 
  io %>% 
  select(-fold, -x) %>% 
  filter(., Condition.Exposure == "prior") %>% 
  # make 3 rows of prior io condition then join each prior with one of the three exposure conditions
  mutate(rows = 3) %>% 
  uncount(rows) %>% 
  mutate(
    match = c("Shift0", "Shift10", "Shift40"), 
    Condition.Exposure = str_c(Condition.Exposure, match)) %>% 
  right_join(
    d.for_analysis %>% 
      filter(Phase == "exposure") %>% 
      nest(data = -c(ParticipantID, Condition.Exposure)) %>% 
      group_by(Condition.Exposure) %>% 
      # get 1 set of exposure trials per condition
      slice_sample(n = 1) %>% 
      unnest(data) %>% 
      reframe(Item.VOT, Item.f0_Mel, vowel_duration, category) %>% 
      rename(VOT = Item.VOT, f0_Mel = Item.f0_Mel, intended.category = category) %>% 
      # nest the cues into one column
      mutate(x = pmap(list(VOT, f0_Mel, vowel_duration), ~ c(..1, ..2, ..3)),
             # create a temporary variable to enable the join between prior io and each set of exposure from each condition
             match = Condition.Exposure) %>% 
      select(c(match, intended.category, x)) %>%  
      nest(x = c(x, intended.category))) %>% 
  select(-match) %>% 
  # get IO categorisation with both decision rules
  mutate(
    categorization.proportional = 
      map2(x, io, ~ 
             get_categorization_from_MVG_ideal_observer(x = .x$x, model = .y, decision_rule = "proportional") %>% 
             filter(category == "/t/")),
    categorization.criterion = 
      map2(x, io, ~ 
             get_categorization_from_MVG_ideal_observer(x = .x$x, model = .y, decision_rule = "criterion") %>% 
             filter(category == "/t/"))) %>% 
  unnest(c(3:5),
         names_sep = "_") %>% 
  select(!ends_with(c("observationID", "_x", "_category"))) %>% 
  # for all items that are not intended /t/ take 1 minus the posterior of /t/
  mutate(
    categorization.proportional_response = ifelse(x_intended.category == "/t/", categorization.proportional_response, 1 - categorization.proportional_response),
    # where criterion rule responds 1 that is a /t/
    categorization.criterion_response = ifelse(categorization.criterion_response == 1, "/t/", "/d/"),
    categorization.criterion_response = ifelse(categorization.criterion_response == x_intended.category, T, F)) %>% 
  group_by(Condition.Exposure) %>% 
  summarise(across(c(categorization.proportional_response, categorization.criterion_response), mean)) %>% 
  bind_rows(
    io %>% 
      select(-fold, -x) %>% 
      filter(Condition.Exposure != "prior") %>% 
      right_join(
        d.for_analysis %>% 
          filter(Phase == "exposure") %>% 
          nest(data = -c(ParticipantID, Condition.Exposure)) %>% 
          group_by(Condition.Exposure) %>% 
          # get 1 set of exposure trials per condition
          slice_sample(n = 1) %>% 
          unnest(data) %>% 
          reframe(Item.VOT, category) %>% 
          rename(x = Item.VOT, intended.category = category) %>% 
          nest(x = c(x, intended.category))) %>% 
      mutate(
        categorization.proportional = map2(
          x, 
          io, 
          ~ get_categorization_from_MVG_ideal_observer(x = .x$x, model = .y, decision_rule = "proportional") %>% filter(category == "/t/")),
        categorization.criterion = map2(
          x, 
          io, 
          ~ get_categorization_from_MVG_ideal_observer(x = .x$x, model = .y, decision_rule = "criterion") %>% filter(category == "/t/"))) %>%
      unnest(c(3:5), names_sep = "_") %>% 
      select(!ends_with(c("observationID", "_x", "_category"))) %>% 
      # for all items that are not intended /t/ take 1 minus the posterior of /t/
      mutate(
        categorization.proportional_response = ifelse(x_intended.category == "/t/", categorization.proportional_response, 1 - categorization.proportional_response),
        # where criterion rule responds 1 that is a /t/
        categorization.criterion_response = ifelse(categorization.criterion_response == 1, "/t/", "/d/"),
        categorization.criterion_response = ifelse(categorization.criterion_response == x_intended.category, T, F)) %>% 
      group_by(Condition.Exposure) %>% 
      summarise(across(c(categorization.proportional_response, categorization.criterion_response), mean))) %>% 
  rename(Condition.io = Condition.Exposure) %>% 
  mutate(Condition.Exposure = factor(rep(c("Shift0", "Shift10", "Shift40"), 2))) %>% 
  pivot_longer(
    cols = starts_with("categorization"),
    names_to = "decision",
    values_to = "accuracy") %>% 
  mutate(
    Condition.io = ifelse(str_detect(Condition.io, "prior"), "prior", Condition.io),
    decision = factor(str_replace(decision, "categorization\\.(.*)_response", "\\1"), levels = c("proportional", "criterion"))) 

# get predictions of the fitted psychometric model
fit_exposure <- readRDS("../models/exposure-standard-priorSD15-0.999.rds")
fit_test <- readRDS("../models/test-standard-priorSD15-0.995.rds")

psych.model_accuracy <- 
  get_pyschometric_accuracy(fit_exposure, d.for_analysis, "exposure", VOT.sd_exposure) %>% 
  bind_rows(
    get_pyschometric_accuracy(fit_test, d.for_analysis, "test", VOT.sd_test)) %>% 
  mutate(decision = factor(decision)) %>%
  ungroup() %>%
  mutate(decision = fct_relevel(decision, "proportional", "criterion")) %>% 
  add_block_labels()

rm(fit_exposure)
```

```{r IO-human-accuracy, fig.width=7.5, fig.height=base.height*2 + 1/2, fig.cap="(ref:IO-human-accuracy)", fig.pos="!ht"}
psych.model_accuracy %>% 
  group_by(decision, Condition.Exposure) %>% 
  arrange(Block, .by_group = T) %>% 
  ggplot(
    aes(x = Block.plot_label, y = median, 
        ymin = lower, ymax = upper, 
        colour = Condition.Exposure, 
        group = decision)) +
  geom_hline(
    data = io.accuracy %>% filter(Condition.io != "prior"),
    mapping = aes(yintercept = accuracy, color = Condition.Exposure),
    linetype = 2,
    linewidth = .8,
    alpha = 0.4,
    show.legend = F) +
  geom_hline(
    data = io.accuracy %>% filter(Condition.io == "prior"),
    mapping = aes(yintercept = accuracy),
    linewidth = .8,
    colour = "#d2d4dc",
    alpha = 0.9,
    inherit.aes = F) +
  geom_pointrange(position = position_dodge2(.5), size = .4) +
  stat_summary(geom = "line", position = position_dodge2(.4)) +
   labs(y = "Accuracy\n(of psychometric model)",
        x = "Block") +
  scale_colour_manual("Condition", values = colours.condition, labels = c("baseline", "+10", "+40"), guide = F) +
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1)) + 
  coord_cartesian(ylim = c(.5, 1)) +
  guides(shape = "none") +
  facet_grid(decision ~ Condition.Exposure,
             labeller = labeller(decision = c("proportional" = "Luce's\ndecision rule", "criterion" = "Criterion\ndecision rule")))
```

If confirmed, premature convergence against stable behavior despite only partial adaptation would challenge prediction (3) of existing distributional learning models. Premature convergence is also unexpected under any other model of adaptive speech perception. In the general discussion, we present an extension to distributional learning models that explains premature convergence, and highlights a striking link between adaptive changes in speech perception and second language learning. As part of that discussion, we present additional models, and entertain methodological artifacts, and analysis confounds that would offer alternative explanation of premature convergence. For now, we summarize one additional property of incremental adaptation suggested by the data.

Further comparison of exposure conditions in Figure \@ref(fig:plot-fit-PSE)C reveals that more extreme exposure conditions---conditions that differed more from listeners' pre-exposure expectations---might have resulted in proportionally less adaptation. Consider the two conditions that are shifted leftwards relative to listeners pre-exposure expectations: in the more extreme baseline condition, participants' PSE had shifted `r d.true_shift %>% filter(Condition.Exposure == "Shift0" & Block == 7) %>% summarise(proportion_shift = percent(proportion_shift)) %>% pull()` by Test 4, compared to `r d.true_shift %>% filter(Condition.Exposure == "Shift10" & Block == 7) %>% summarise(proportion_shift = percent(proportion_shift)) %>% pull()` in the less extreme +10 condition (`r str_extract(get_bf(fit_test_nested_within_condition_and_block, str_c("mu2_IpasteCondition.ExposureBlocksepEQxShift0x7 / ", predictedPSE_0, " < mu2_IpasteCondition.ExposureBlocksepEQxShift10x7 /", predictedPSE_10)), "BF = \\d")`). That is, while the absolute changes in listeners' behavior exhibit the pattern expected under predictions (1) and (2b), participants additionally exhibited a pattern of **shrinkage** that is *not* obviously predicted by distributional learning models. 

Finally, there is evidence that this shrinkage is asymmetric. The predicted PSEs of an idealized learner for the baseline and +40 conditions are shifted by about the same amount relative to listeners' pre-exposure PSE in Test 1 (+21ms and -18ms VOT, respectively). However, the degree to which listeners converged against these predicted PSEs differed substantially between the two conditions, with cumulative adaptation proceeding almost twice as far in the rightward-shifted +40 condition (in Test 4: `r d.true_shift %>% filter(Condition.Exposure == "Shift40" & Block == 7) %>% summarise(proportion_shift = percent(proportion_shift)) %>% pull()` towards idealized PSE) compared to the leftward-shifted baseline condition (`r d.true_shift %>% filter(Condition.Exposure == "Shift0" & Block == 7) %>% summarise(proportion_shift = percent(proportion_shift)) %>% pull()`; `r str_extract(get_bf(fit_test_nested_within_condition_and_block, str_c("mu2_IpasteCondition.ExposureBlocksepEQxShift0x7 / ", predictedPSE_0, "< mu2_IpasteCondition.ExposureBlocksepEQxShift40x7 /", predictedPSE_40)), "BF = \\d{2}.\\d")`). Taking this **asymmetric shrinkage** at face value (for now), it suggests <!-- revise wording to reflect strength of evidence: anecdotal, moderate, etc. --> that participants adapted to a lesser degree to leftward shifts along the VOT continuum. We return to this point in the general discussion, where we also summarize other recent studies that have found similar asymmetric shrinkage [@kleinschmidt-jaeger2016; @kleinschmidt2020]. Before we do so, we briefly summarize the effects of repeated testing.

## Effects of repeated testing

```{r calculate-test-surprisal}
d.test_surprisal <-
  d.for_analysis %>%
  filter(Phase == "test") %>%
  distinct(Item.VOT) %>%
  cross_join(io %>% select(-c(x, fold))) %>%
  filter(Condition.Exposure %in% c("Shift0", "Shift10", "Shift40")) %>%
  # For each test stimulus, calculate the sum densities of /d/ and /t/ over
  # that VOT (i.e., the density of the marginal VOT distribution)
  mutate(
    density = map2_dbl(
      Item.VOT,
      io,
      function(x, y)
        pmap(
          list(y$mu, y$Sigma, y$Sigma_noise),
          ~ dmvnorm(x, ..1, ..2 + ..3)) %>%
        reduce(`+`)),
    surprisal = -log2(density))


# # Visualize marginal density
# d.test_surprisal %>%
#   ggplot(aes(x = Item.VOT, y = density, color = Condition.Exposure)) +
#       geom_line()
#
# # Visualize surprisal
# d.test_surprisal %>%
#   ggplot(aes(x = Item.VOT, y = surprisal, color = Condition.Exposure)) +
#       geom_line()

d.test_surprisal %<>%
  group_by(Condition.Exposure) %>%
  summarise(across(surprisal, list(mean = mean, sd = sd)))
```

Some models of adaptive perception predict that exposure to uniformly distributed test tokens will reduce the effect of preceding exposure [@kleinschmidt-jaeger2015; for relevant discussion, see also @lancia-winter2013]. In line with these theories, there is evidence that the effects of exposure reduced from Test 4 to Test 6 (see Tables \@ref(tab:hypothesis-table-simple-effects-condition) and \@ref(tab:hypothesis-table-simple-effects-block)). In Table \@ref(tab:hypothesis-table-simple-effects-block), this is evident in a reversal of the direction of the block-to-block changes for Tests 5-6, compared to Tests 1-4. For the +40 exposure condition, these block-to-block changes went from rightward shifts in Tests 1-4 to leftward shifts in Tests 5-6 (BF = 10.4). For the baseline condition, block-to-block changes went from leftward to rightward shifts (BF = 7.3). The only exposure condition for which no clear reversal was observed is the +10 condition (BF = 1.3). Two factors likely contributed to this. First, this condition exhibited the smallest exposure effects, limiting the power to detect a reversal of those effects. Second, the +10 condition is also the condition, for which the marginal distribution of VOT during test blocks (mean = 35.8 ms, SD = 22.2 ms) most closely resembled the distribution during exposure (mean = 36.5, SD = 25.9), compared to the baseline (mean = 26.5 ms) or +40 condition (mean = 66.5 ms; exposure SDs were identical across conditions).

As a consequence of repeated testing, exposure effects were substantially smaller in Test 6 than in Test 4 (see Table \@ref(tab:hypothesis-table-simple-effects-condition): while the effects of the +40 condition relative to the other two exposure conditions were still credible even in Test 6 (BFs > 24), this was no longer the case for the effect of the +10 condition relative to the baseline condition (BF = 1.6). This pattern of results replicates previous findings from LGPL [@scharenborg-janse2013; @giovannone-theodore2021; @cummings-theodore2023; @liu-jaeger2018; @liu-jaeger2019; @zheng-samuel2023], and extends them to distributional learning paradigms [see also @colby2018; @kleinschmidt2020]. One important methodological consequence is that longer test phases do not necessarily increase the statistical power to detect effects of adaptation [unless analyses account for the effects of repeated testing, as done in, e.g., @liu-jaeger2018]. <!-- TO DO: add @cummings2024 once made available online (pen in mouth paper --> Analyses that average across all test tokens---as remains the norm---are bound to systematically underestimate the true adaptivity of human speech perception.<!-- ^[@kraljic-samuel2006 is sometimes cited as finding LGPL exposure effects even after 480 test trials over a uniform test continuum. This is, however, misleading. Kraljic and Samuel used four *different* uniform test continua over two different phonetic contrasts (/b/-/p/ and /d/-/t/). Each test session consisted of 10 randomized repetitions of 6 test trials. Kraljic and Samuel never tested (or made any claims about) whether exposure effects were still detectable during the 10th repetition. Rather they report *average* effects across the 10 repetitions (like other LGPL studies), which is perfectly compatible with the hypothesis that repeated testing reduces the effects of exposure [see @liu-jaeger2018].] -->


```{r}
rm(fit_test_nested_within_condition_and_block)
```

