<!-- Do NOT knit this document. It is part of a larger document. Instead knit the main document (my-apa-formatted-article) -->

# General discussion
Over the last 20+ years, landmark studies in adaptive speech perception have demonstrated that listeners' interpretation of speech is not static. Instead, it can change with recent exposure, accommodating differences in pronunciation across talkers. Even a single session of exposure can be sufficient to substantially reduce losses in processing speed or accuracy that listeners might initially experience when exposed to unfamiliar speech patterns. Research on accent adaptation [AA, @eisner2013; @schertz2016; @xie2017] and perceptual recalibration [VGPL/LGPL, @eisner-mcqueen2005; @kraljic-samuel2006; @kurumada2018; @norris2003; @reinisch-holt2014; @vroomen2007] has found that this flexibility is achieved through changes in listeners' *categorization functions*---the mapping from acoustic or phonetic cues to the phonological categories that form the input to spoken language understanding. These changes in listeners' categorization functions seem to depend on the statistics of the exposure input in ways qualitatively expected under the hypothesis that adaptive speech perception involves distributional learning [DL, @clayards2008; @idemaru-holt2011; @idemaru-holt2020; @kleinschmidt-jaeger2015; @kleinschmidt2020; @mcmurray-jongman2011; @nixon2016; @theodore-monto2019].

Despite these advances, little is known about the incremental unfolding of adaptive speech perception [@xie2023; @OTHERS]. In particular, it remains unclear how listeners' categorization functions change incrementally with exposure, and how those changes depend on the amount and cumulative distribution of phonetic evidence in the input listeners have experienced so far. The present study modified a distributional learning paradigm to investigate these questions. We found that listeners' categorization functions changed incrementally with exposure, and that the direction and magnitude of that change depended on listeners' prior expectations based on previously experienced speech input from other talkers (Prediction 1 from the introduction), and both the amount and distribution of phonetic evidence in the exposure input from the unfamiliar talker (Predictions 2a and 2b, respectively). Our analyses also suggest two properties of incremental adaptation that go beyond these qualitative predictions. First, we found that the largest changes in listeners' categorization functions occurred during the initial moments of exposure. And, second, while exposure rapidly caused substantial changes in listeners' categorization functions, listeners did *not* fully adapt to the unfamiliar talker's speech patterns.

In the remainder of this article, we discuss these findings in more detail, and consider their implications for theories of adaptive speech perception. We begin with the question of how early changes in listeners' categorization function can be detected. Then we turn to predictions (1) and (2a,b), before discussing potential constraints on adaptive speech perception that might prevent listeners from fully adapting to unfamiliar speech patterns. We close by considering limitations of the present study and how future work can overcome them.

## How quickly can listeners adapt their speech perception?
We found significant shifts in listeners' categorization function even after the briefest exposure tested. Exposure to 24 tokens each of shifted /d/ and /t/ was sufficient to significantly change how listeners interpreted subsequent inputs. Of note, only half of these exposure tokens labeled the intended category, the other half did not. Even when trials were labeled, labeling was indirect rather than through explicit feedback: on labeled trials, the two response options listeners saw both had the same onset stop (e.g., "din" and "dill"). Previous distributional learning studies have assessed exposure effects after *much* longer exposures, typically hundreds of trials [e.g., 192 trials in @harmon2019; 200 in @idemaru-holt2011; 228 in @clayards2008; @kleinschmidt-jaeger2016; 236 in @theodore-monto2019; 456 in @nixon2016]. The present results demonstrate that a fraction of the amount of exposure employed in previous studies is sufficient to elicit changes in listeners' categorization behavior.

This result bears on ongoing discussions as to whether the type of adaptive changes observed in AA and LGPL/VGPL paradigms could plausibly originate in the same mechanisms as those observed in DL paradigms like in the present study [@bradlow-bent2008; @baese-berk2018; @zheng-samuel2020; @xie2023]. A number of studies have now demonstrated improvements in the speed or accuracy of speech perception after relatively short exposure to an unfamiliar L2-accented talker [@bradlow2023; @clarke-garrett2004; @xie2018; @xie2021jep]. For example, Xie and colleagues [-@xie2018] found improvements in both the speed and accuracy of cross-modal priming after exposure to only 18 sentences from an L2-accented talker---the shortest tested exposure we are aware of [see also @clarke-garrett2004]. Other AA work has directly assessed changes in the categorization function. For example, @xie2017 found changes in listeners categorization behavior after exposure to only 30 critical L2-accented words. The directionality of these changes was consistent with distributional learning accounts of adaptive speech perception. Together with evidence from additional experiments, Xie and colleagues concluded that "listeners dynamically update their own cue-weighting functions during rapid phonetic adaptation to foreign accents, and critically over much shorter time span[s] than shown in previous studies of second language phoneme learning". By demonstrating that DL paradigms can elicit qualitatively similar changes with similarly little exposure, the present study lends further plausibility the hypothesis that these changes are driven by the same underlying mechanisms.

Some experiments on LGPL/VGPL have demonstrated effects after even less exposure, with detectable changes in listeners' categorization responses after as few as 2-4 exposures to visually or lexically labeled phonetically tokens [@cummings-theodore2023; @kleinschmidt-jaeger2011; @kleinschmidt-jaeger2012; @liu-jaeger2018; @liu-jaeger2019; @vroomen2007]. In comparing findings across paradigms, future work should keep in mind that DL and LGPL/VGPL paradigms differ in the amount of information conveyed by each exposure token. LGPL/VGPL paradigms typically employ exposure stimuli that a) are labeled and b) auditorily maximally ambiguous---falling between the two categories that the experiment focuses on. Distributional learning accounts predict that such stimuli should be highly informative, leading to comparatively large changes in categorization behavior. This is in line with recent findings: when stimuli in LGPL/VGPL experiments are shifted less than to the point of maximal auditory ambiguity, listeners exhibit smaller shifts in categorization behavior [@babel2019; @kleinschmidt-jaeger2012; @tzeng2021; for a numerical but non-significant replication, see also @cummings-theodore2023]. In contrast, DL paradigms a) typically employ only unlabeled stimuli [@clayards2008] or a mixture of unlabeled and labeled stimuli [e.g., @kleinschmidt-jaeger2016 and the present paradigm], and b) reflect a *distribution* of phonetic properties---ranging from more to less expected under listeners' prior expectations. This makes the speech inputs in DL paradigms more similar to what one would expect during exposure to natural accents and other cross-talker differences. But it also makes exposure tokens in DL experiments, on average, considerably less informative than in an LGPL/VGPL experiment. Future work that aims to compare the speed of adaptive speech perception across these two paradigms should thus do so *relative to the amount of information conveyed by each exposure*.

<!-- TO DO: *could* add a figure here showing how the PSE would be predicted to change for our experiment under a certain kappa, nu combination, and compare that against PSE changes for the same data when all /d/ stimuli are set to *prior* PSE value and all /t/ stimuli are set to prior /t/ (or vice versa) -->

## Incremental adaptation based on the amount and distribution of phonetic evidence (Predictions 1 and 2a,b)
To the best of our knowledge, the present study is the first to assess the joint effects of prior and recent exposure gradiently unfold with increasing exposure, predictions (10 and (2a,b) described in the introduction. While most contemporary theories of adaptive speech perception share these qualitative predictions, relatively few experiments have investigated how listeners' categorization functions change with continued exposure to a phonetic distribution. We briefly review potential reasons for this before we turn to our results.

### Previous work
The majority of studies on AA, for example, continue to employ the popular single exposure-test design without incremental testing. In this design, different groups of listeners are exposed to different speech input. The effects of that exposure are then assessed by a single post-exposure test, which typically is held identical across the groups. This design is well suited to assess the qualitative effects of exposure to different types of speech input---for example, whether familiarization with a talker's speech generally leads to improved speech recognition---but it sheds little light on *how* the phonetic properties of the speech input shape changes in listeners' perception. If repeated testing has been been employed in research on AA, it tended to be after considerably longer exposure [e.g., testing on different days, @eisner-mcqueen2006; but see @xie-kurumada2024]. The same largely holds for previous DL studies, which either employed a single batch test following exposure [e.g., @clayards2008; @idemaru-holt2020; @nixon2016; @schertz2016] or repeated testing over the course of multiple days [e.g., @mcclelland1999].

One reason for the relative scarcity of studies on incremental changes during the early stages of exposure might originate in the focus of earlier research. As mentioned in the introduction, the pioneering generation of studies focused on the question of *whether* adaptation is observed, and *under what conditions* [for review, see @cummings-theodore2023]. Now that the existence of adaptive changes in speech perception is no longer in question, paradigms that inform and constrain the functional relation between exposure inputs and changes in listeners' perception have gained in importance. At the minimum, this includes a clearer understanding of how the amount and statistics of the exposure input drive changes in speech perception. Stronger tests of competing computational models will require such data. In a recent review of the field, @xie2023 demonstrated that the signature findings of some of the most popular paradigms in adaptive speech perception do not distinguish between radically different theoretical accounts: qualitative improvements in speech recognition can be explained by mechanisms ranging from early pre-linguistic perceptual normalization, changes in the representations of phonetic categories, or upstream changes in decision-making. Other recent reviews, too, have highlighted the need to develop stronger theories, sufficiently specific to be implemented as predictive computational models [@guest-martin2021; @kleinschmidt-jaeger2015; @norris2017; @yarkoni-westfall2017]. Effective comparisons of these theories will require quantitative data sets that constrain the way in which listeners' categorization behavior changes depending on the amount and nature of the input.

A second reason for the relative scarcity of research on incremental changes in speech perception might be that repeated testing comes with its own unique challenges. For example, if test tokens are sampled from natural accents---the most common approach in AA research---these tokens can themselves contain information about the target accent, thereby masking exposure effects [for a particularly clear demonstration, see @xie-kurumada2024]. Even for paradigms that employ synthesized stimuli, or otherwise control the informativity of test tokens [e.g., @chodroff-wilson2020], repeated testing can come with challenges. There is now evidence that repeated testing over unlabeled test continua that uniformly span the relevant phonetic space can reduce the effects of exposure [@scharenborg-janse2013; @cummings-theodore2023; @zheng-samuel2023; @liu-jaeger2018; @liu-jaeger2019; @giovannone-theodore2021; @tzeng2021]. This is compatible with some distributional learning accounts of adaptive speech perception [for discussion, see @kleinschmidt-jaeger2015]: if the same learning mechanisms that operate during labeled and unlabeled exposure trials continue to operate during unlabeled test trials, the unexpected uniform distribution over the VOT continuum during test blocks should begin to undo the effects of preceding exposure.^[The specific predictions depend on the---as of yet unknown---way listeners adapt to unlabeled speech inputs [for an initial comparisons of several candidate models for unsupervised adaptation, see @yan-jaeger2018]. Additionally, it is possible that adaptation to unlabeled inputs involves different mechanisms than adaptation to labeled trials.] In the present study, we replicated this effect of repeated testing for the final three test blocks. For research on incremental adaptation this implies a methodological tension between longer test phases, which provide researchers with more observations, and the risk of diminishing effects [see also @liu-jaeger2018].

### The present study
In the present study, we used a repeated exposure-test design to assess incremental changes in listeners' categorization functions from pre-exposure onward. We found that that listeners' categorization function shifted with increasing exposure to shifted phonetic distributions. The direction and magnitude of these shifts developed gradiently and were qualitatively consistent with the predictions of distributional learning accounts [e.g., @apfelbaum-mcmurray2015; @harmon2019; @johnson1997; @kleinschmidt-jaeger2015; @magnuson2020; @assmann-nearey2007; @sohoglu-davis2016; @xie2023]. This was the case both for shifts in listeners' categorization function relative to other exposure conditions (Table \@ref(tab:hypothesis-table-simple-effects-condition)) and for shifts in listeners' categorization function relative to their pre-exposure categorization function (Table \@ref(tab:hypothesis-table-simple-effects-block)). In both cases, every single test we conducted went in the direction predicted by distributional learning theories. We elaborate on these findings.

### Prediction 1: The starting point of adaptation depends on listeners' prior experience
The inclusion of a pre-exposure test made it possible to assess prediction (1)---that shifts in listeners' categorization function should depend on how the exposure distributions *relative to listeners' relevant prior experiences*. This prediction received support both from the fact that listeners' responses prior to exposure were well approximated based on a database of /d/ and /t/ productions, and from the fact that the direction of changes in listeners' PSEs *relative to pre-exposure* was predicted by the direction of the shift in /d/ and /t/ distributions relative to their distributions in prior experience (Figure \@ref(fig:exposure-means-database-matrix-plot)).

 + caveat that prior predictions depend on database.  
  + we chose it to match the genre. if one was to, for example, use a database of connected speech, one would predict a different range of PSE. One possible interpretation for this is that listeners have not just one set of prior expectations but rather a whole set, and that they are able to recognize the speech style/register a talker is using.
  + however, this should be interpreted with caution. the database we chose was relative small. We tried to provide a sense of the uncertainty one should maintain about these estimates (CV). Still future work should user larger databases. a challenge in this endevour will be to determine *how* listeners figure out *what* previous experience is relevant to the current input.

### Prediction 2a: Adaptation increases with the amount of exposure
Prediction (2a)---that the magnitude of changes in listeners' categorization function should gradiently increase with the *amount* of exposure---received support from the comparisons across blocks: increasing exposure consistently yielded additional shifts in listeners' PSE. This replicates in a DL paradigm recent findings from VGPL/LGPL experiments [@cummings-theodore2023; @kleinschmidt-jaeger2012; @liu-jaeger2018; @liu-jaeger2019; @vroomen2007]. In a particularly careful study, Cummings and Theodore compared shifts in categorization function between groups of listeners after exposure to 1, 4, 10, or 20 lexically labeled shifted /s/ or /`r linguisticsdown::cond_cmpl("ʃ")`/ tokens (each matched by an equal number of unshifted tokens from the opposite category). Shifts in listeners' categorization functions increased with the number of exposure to tokens, and did so significantly for all comparisons, except but the comparison between 4 and 1 exposures. @vroomen2007 found similarly increasing shifts in categorization functions *within* participants, comparing the effects of 1, 2, 4, ..., 32 exposures to visually labeled shifted tokens [see also @kleinschmidt-jaeger2012].^[With increasing exposure, the direction of shift begins to reverse [returning to baseline after 128-256 exposures, @kleinschmidt-jaeger2011; @vroomen2007] and can even change directional altogether, depending on the degree of the shift [@kleinschmidt-jaeger2012]. Later work showed that this reversal is accounted for by distributional learning models, and caused by the fact that the VGPL paradigm presents the exact same stimulus on each exposure trial [@kleinschmidt-jaeger2015].] The present study demonstrated similarly gradient effects with increasing exposure to a mixture of labeled and unlabeled exposure tokens that were randomly sampled from a *distribution* of phonetic tokens, thereby more closely resembling the situation listeners would experience during everyday speech perception. This aspect of our results thus adds to a growing number of similarities in the findings between LGPL/VGPL and DL paradigms, as expected under the hypothesis that changes observed in both paradigms originate in the same underlying mechanisms [see discussions in @kleinschmidt2015; @zheng-samuel2020; @xie2023].

### Prediction 2a: Adaptation depends on the phonetic distribution in the exposure input
Finally, prediction (2b)---that the direction and magnitude of changes in listeners' categorization function should depend on the *phonetic distribution* of the exposure input---received support from the within-block comparisons across exposure conditions: the direction of the shift of the /d/ and /t/ category means in the exposure input correctly predicted the relative ordering of listeners' PSEs in Test 1-4, and shifts in category means of larger magnitude (+40 vs. baseline compared to +10 vs. baseline) yielded larger shifts in listeners' PSE. This extends the findings from previous DL studies to demonstrate gradient incremental adaptation towards the exposure distribution.

# Beyond qualitative tests
Going beyond this qualitative test of prediction (2b), two of tests we conducted constitute an initial step towards stronger quantitative tests.

## First fast, then slow: diminishing returns of exposure
One test found suggestive evidence that the speed of incremental changes in listeners' PSE decreased with increasing exposure: the same amount and distribution of phonetic evidence yielded smaller *additional* changes in listeners' PSE, the more exposure blocks listeners had already experienced. To the best of our knowledge, this is the first study to report this pattern for adaptive speech perception. A similar pattern of 'diminishing returns' is, however, also indirectly evident in at least one other recent studies. Kleinschmidt [-@kleinschmidt2020, Experiment 3] re-analyzes several DL experiments originally presented in @kleinschmidt-jaeger2016. In these experiments, which very much informed the present study, different groups of participants were exposed to VOT distributions that were shifted to different degrees. Unlike the present study, the experiments by Kleinschmidt and Jaeger were not designed to detect incremental changes in listeners' categorization function: they lacked incremental tests over identical VOTs, allowing trial order to confound VOT [see discussion in @kleinschmidt2020]. Still, Kleinschmidt's Figure 9, partially replotted here as Figure XXX, is compatible with the hypothesis that adaptation initially proceeded quickly, and then increasingly more slowly. This is particularly evident for the more extreme exposure conditions, though the data are (even) more noisy than for the present study.

<!-- TO DO: add description of Dave's study; incl. absence of incremental testing, and the way it was 'hacked'/approximated in the modeling in section XXX of his ms. NOTE: this is NOT about shrinkage. it's about diminishing returns -->

Such diminishing returns of exposure are explicitly predicted only by some distributional learning models [e.g., error-based models, @davis-sohoglu2020; @sohoglu-davis2016; Bayesian ideal adaptors, @kleinschmidt-jaeger2015; @kleinschmidt-jaeger2016]: prior to adaptation, an idealized listener (horizontal gray ribbons in Figure \@ref(fig:plot-fit-PSE)C) would, on average, experience larger prediction errors (mean surprisal per exposure input in baseline condition ${\rm E}[-\log_2 p(category | VOT_i, f0_i, vowel\ duration_i)] =$ `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift0", IO.Type == "prior") %>% pull(surprisal_mean) %>% round(., 2)` bits; +10 condition = `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift10", IO.Type == "prior") %>% pull(surprisal_mean) %>% round(., 2)` bits; +40 condition = `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift40", IO.Type == "prior") %>% pull(surprisal_mean) %>% round(., 2)` bits). As listeners converge towards the distribution of /d/ and /t/ in the exposure condition, they should experience increasingly smaller prediction errors processing the exposure tokens. An idealized learner that has fully converged against the exposure distributions (colored lines in Figure \@ref(fig:plot-fit-PSE)C) would, on average, experience only `r d.exposure_surprisal %>% filter(Condition.Exposure == IO.Type) %>% pull(surprisal_mean) %>% mean(.) %>% round(., 2)` bits of surprisal per exposure input. Models of adaptive speech perception that predict adaptation to be a positive monotonic function of the prediction error, thus offer a qualitative explanation for the the diminishing returns of exposure observed in the present study. If this pattern is replicated in future work, this would raise the question as to whether similar predictions follow from other distributional learning accounts [e.g., C-CuRE normalization, @mcmurray-jongman2011; exemplar models, @johnson1997; DNNs, @magnuson2020]. If, on the other hand, future tests fail to replicate these findings, this would be a serious problem for models that predict adaptation to be sensitive to the prediction error.

If replicated, it would be tempting to interpret these 'diminishing returns' as evidence that listeners have converged against the exposure distributions in the input---i.e., that adaptation has successfully completed. In particular, the hypothesis tests in Table \@ref(tab:hypothesis-table-simple-effects-block) suggested that there was at best anecdotal evidence that the final exposure block had any additional effect---including in the two conditions with the largest shifts. However, our findings suggests that this does *not* mean that listeners had successfully learned all required information about the unfamiliar talker: compared to PSEs predicted for idealized learners that have fully learned the exposure distributions, listeners' exhibited substantially smaller shifts in PSEs. We discuss possible explanations of this observation next.

## Constraints on (the early moments of) adaptive speech perception?

 Why constrained? Why asymmetric?

 + methodological issues that would mean ideal adaptor might still be compatible with data
  + regularizing priors pulling ...?
  + some subjects not doing the task and we're not modeling that (zero adaptors not being modeled separately might pull more extreme conditions more to the mean)
  + failure to model learning during test: intermittend test undoing effects even within 12 trials? if so, that's interesting. future work can test this by ...

 + ideal adaptor but with modification:
  + talkers are not the relevant context. e.g., could some form of moving window with historical decay explain the findings? On the one hand if the moving window is very small, that would not explain why we do see some *cumulative* changes across blocks (window must be at least 48 + 12 = 60 trials). on the other hand, the qualitative changes in the PSEs and slopes suggest that 12 trials can be enough to change some aspects of the categorisation function. it's thus *possible* that something that ways recent input much more strongly but also considers less recent input beyond 48 trials might explain the overall pattern.

 + DL might be correct but ideal adaptor *model* as used here might be wrong
  + goal of learning (loss function) is not approximation of distributions (i.e., maximizing likelihood of joint distribution of cues and categories) but e.g., maximization of accuracy under a choice rule.
   + COULD (but maybe not?) discuss potential that observed adaptation maximizes accuracy under the choice rule. use psychometric function fit during unlabeled exposure trials to calculate *accuracy* (not likelihood) on labeled trials under criterion and under proportional matching decision rules. compare against accuracy if ideal observers categorization functions are used instead.
  + difficulty extrapolating to unobserved VOTs? [@kleinschmidt2020, p. 27].
  + limitations in incremental adaptation (that might potentially be overcome with longer-term exposure, cf. discussion in xie2023): e.g., reweighting mixture of previously learned talker-models (can't go beyond previous experience).


All of these tests were facilitated by the use of incremental testing, which thereby facilitated stronger---though still qualitative---tests of theories of adaptive speech perception. The use of incremental testing also revealed

## Constraints on rapid adaptation?
To further quantify these visual intuitions, we fit a distributional learning model to listeners' data. This allows us to compare changes in listeners' behavior to those expected under an ideal---as opposed to ideal*ized*---learner.^[Following @qian2016, we use the term *idealized* learner for models that are given privileged access to information that listeners have to infer (e.g., the correct exposure statistics of the experiment), and the term *ideal* learner for models that implement a theory of ideal information integration [for a list of assumptions made by the ideal adaptor, see appendix of @kleinschmidt-jaeger2015].] Specifically, we used an ideal adaptor model [@kleinschmidt-jaeger2015; @kleinschmidt-jaeger2016]. <!-- TO DO: fill in --> Univariate instances of this model have previously achieved high accuracy in modeling changes in listeners categorization functions after LGPL/VGPL [e.g., for /s/-/`r linguisticsdown::cond_cmpl("ʃ")`/, @cummings-theodore2023; /b/-/d/, @kleinschmidt-jaeger2011; @kleinschmidt-jaeger2012] or DL exposure [e.g., for /b/-/p/, @kleinschmidt-jaeger2016; @kleinschmidt2020; /g/-/k/, @theodore-monto2019]. We extended the model to multivariate categories over VOT, f0, and vowel duration, and fit it to listeners' responses in our experiment (for details, see SI \@ref(sec:XXX)). <!-- TO DO: fill in more info here. Perhaps we should take a model that uses, as priors, the mu_0 and Sigma_0 from the gray ribbon in the result figure -->
Unlike the psychometric model we used for analysis, a distributional learning model specifies how information is integrated across exposure trials. This substantially decreases the degrees of freedom the model has to fit listeners' responses. For example, the ideal adaptor we fit to listeners' responses in the first four test blocks has 19 parameters, compared to 25 fixed and 337 random effect parameters for a psychometric mixed-effect model fit to the same data. <!-- TO DO: calculate these numbers. for psychometric model, it's probably easiest to just count fixed and random effect if model is fit to only first four test data [doesn't have to converge so you can just take 100 samples] --> In the ideal adaptor, listeners' responses during different test blocks and across different exposure conditions jointly constrain the parameters of model (due to assumption of ideal information integration). For the present data, the ideal adaptor model predicts the proportion of listeners' "t"-responses with an $R^2 > .96X$.<!-- TO DO: fill in--> We leave further analysis of this model to future work. Here we focus on what this model can tell us about the incremental build-up of adaptation.

Figure \@ref(fig:plot-IA-human-PSE) shows the PSEs inferred from the ideal adaptor, along with listeners' PSEs (the same as in Figure \@ref(fig:plot-fit-PSE)C). Notably, the ideal adaptor *under*-predicts changes in listeners' PSEs in early blocks---a tendency that is most evident for the exposure conditions with more extreme shifts (baseline and +40). Similarly, though less obvious, the ideal adaptor *over*-predicts listeners' PSE in later blocks (see the baseline and +40 conditions in Test 4). To understand this pattern, it is helpful to recognize that rational models of distributional learning describe listeners' expectations after exposure as the weighted mean of prior expectations and the observed exposure statistics [@kleinschmidt-jaeger2015]. This means that the speed with which a learner adapts at the start of exposure is expected to be predictive of how fast a learner adapts after additional exposure. Put differently, ideal adaptors constrain the rate at which changes in PSEs can change across exposure trials. For the present data, this means that an ideal adaptor can *either* explain the fact that listeners' PSEs changed rapidly at the start of exposure *or* explain the fact that there seem to be little to no changes in listeners' PSE after the second exposure block. But it cannot explain both aspects of our data.  

```{r fit-IA-inferred-VOT-f0}
# Make a data frame that splits the entire data in unique exposure-test combinations.
# This data frame will be used for adaptive changes in normalization and category
# representations.
d_for_ASP <-
  d.for_analysis %>%
  ungroup() %>%
  # Exclude catch trials and the final two test blocks since we expect unlearning during those blocks
  # (which is not modeled)
  filter(
    !(Phase == "exposure" & is.na(Item.ExpectedResponse.Voicing)),
    as.numeric(Block) <= 7) %>%
  # Use F0 as intended by generation script. This makes fitting more computationally feasible
  # (336 instead of 1001 unique combinations of group & test locations), and also removes the
  # potential that annotation / measurement error perturb the f0 values. It does, however,
  # make the assumption that f0 is the same across the three minimal pairs. While likely wrong,
  # this assumption better aligns with the fact that the model doesn't have a way to keep track
  # of any lexical, phonological, or phonetic context effects on VOT (and thus no way to predict
  # any potential differences between minimal pairs).
  # rename(VOT = Item.VOT, f0_Mel = Item.F0_target_for_generation_script) %>%
  # Create new condition variable that uniquely identifies what participants have seen at any of the
  # tests and a new block variable that identifies the type and order of blocks. Also create two
  # convenience variables, x (stimulus) and Response.Category (to make matching with the category
  # column of ideal observers/adaptors more straightforward).
  mutate(
    Condition.for_ASP = paste0(Condition.Exposure, List.ExposureBlockOrder, List.ExposureMaterials),
    Block.for_ASP = paste0(Phase, Block),
    x = map2(VOT, f0_Mel, ~ c(.x, .y)),
    Response.Category = factor(ifelse(Response.Voicing == "voiced", "/d/", "/t/"), levels = c("/d/", "/t/")),
    Item.ExpectedResponse.Category = factor(ifelse(Item.ExpectedResponse.Voicing == "voiced", "/d/", "/t/"), levels = c("/d/", "/t/"))) %>%
  slice_into_unique_exposure_test(
    condition_var = "Condition.for_ASP",
    block_var = "Block.for_ASP",
    block_order = c("test1", "exposure2", "test3", "exposure4", "test5", "exposure6", "test7")) %>%
  mutate(ExposureGroup = factor(ExposureGroup)) %>%
  select(
    ExposureGroup, Phase, ParticipantID, Block, Trial, x,
    VOT, f0_Mel, vowel_duration, Response.Category, Item.ExpectedResponse.Category)

# fit ideal adaptor with uninformative priors
m_IA_inferred.VOT_f0_vowelduration <-
  infer_prior_beliefs(
    exposure = d_for_ASP %>% filter(Phase == "exposure"),
    test = d_for_ASP %>% filter(Phase == "test"),
    cues = c("VOT", "f0_Mel", "vowel_duration"),  
    category = "Item.ExpectedResponse.Category",
    response = "Response.Category",
    group = "ParticipantID",
    group.unique = "ExposureGroup",
    center.observations = T,   
    scale.observations = T,
    pca.observations = F,
    lapse_rate = NULL,
    mu_0 = NULL,
    Sigma_0 = NULL,
    sample = T,
    file = "../models/IBBU_VOT+f0+vowelduration_uninformative priors.RDS",
    warmup = 3000, iter = 4000,
    control = list(adapt_delta = .995, max_treedepth = 15))
```



```{r}
groups <- get_group_levels_from_stanfit(m_IA_inferred.VOT_f0_vowelduration)
# get logistic model predictions of the IA's categorisation
# then prepare the data for plotting
d.IA_predicted_PSE <-
  if (file.exists("../models/d.IA_predicted_PSE.csv")) {
    read_csv("../models/d.IA_predicted_PSE.csv", show_col_types = F)
  } else {
    get_logistic_parameters_from_IBBU(m_IA_inferred.VOT_f0, groups = groups) %>%
      group_by(group) %>%
      median_hdci(PSE) %>%
      # duplicate the no-exposure row; 1 for each condition
      mutate(rows = ifelse(group == "no exposure", 3, 1)) %>%
      uncount(rows) %>%
      mutate(
        Condition.Exposure = factor(c(rep("Shift0", 3), rep("Shift10", 3), rep("Shift40", 3), c("Shift0", "Shift10", "Shift40"))),
        Block = ifelse(group != "no exposure", str_replace(group, ".*(\\d)", "\\1"), group),
        Block = factor(ifelse(group == "no exposure", 1, Block))) %>%
      relabel_blocks()

    write_csv(d.IA_predicted_PSE, "../models/d.IA_predicted_PSE.csv")
  }
```

(ref:plot-IA-human-PSE) Comparison of shifts in listener PSEs over the first four test blocks against that expected by an ideal adaptor fit to listeners' responses (a distributional learning model that describes ideal information integration). Ribbons indicate the 95% highest posterior density interval.

```{r plot-IA-human-PSE, fig.width=6.5, fig.height=3.5, fig.cap="(ref:plot-IA-human-PSE)"}
d.estimates %>%
  filter(Block %in% c(1:7)) %>%
  ggplot(
    aes(
      x = Block.plot_label, y = PSE,
      ymin = PSE.lower, ymax = PSE.upper,
      colour = Condition.Exposure, group = Condition.Exposure)) +
  geom_rect(
    data = tibble(
      xmin = c(seq(0.5, 6.5, 2)),
      xmax = c(seq(1.5, 7.5, 2))),
    mapping = aes(xmin = xmin, xmax = xmax),
    ymin = -Inf, ymax = Inf, fill = "grey", alpha = .1, inherit.aes = F) +
  geom_point(position = position_dodge(.3), size = 1.8) +
  geom_linerange(linewidth = .6, position = position_dodge(.3), alpha = .5) +
  stat_summary(geom = "line", position = position_dodge(.3)) +
  geom_ribbon(
    data = d.IA_predicted_PSE,
    mapping = aes(x = Block.plot_label, ymin = .lower, ymax = .upper, fill = Condition.Exposure, group = Condition.Exposure),
    alpha = .1,
    position = position_dodge(.3),
    inherit.aes = F) +
  geom_linerange(linewidth = .6, position = position_dodge(.3), alpha = .5) +
  geom_hline(
    data = d.io.categorization %>% filter(Condition.Exposure %in% c("Shift0", "Shift10", "Shift40")),
    mapping = aes(yintercept = PSE, color = Condition.Exposure, group = Condition.Exposure),
    linetype = 2,
    linewidth = .8,
    alpha = 0.4) +
  geom_text(
    data = d.true_shift %>%
      filter(Phase == "exposure") %>%
      group_by(Condition.Exposure, true_shift) %>%
      summarise(),
    mapping = aes(
      x = 8, y = c(25, 35, 65),
      label = paste(round(true_shift), "(100%)"),
      colour = Condition.Exposure),
    colour = colours.condition,
    fontface = "bold",
    size = 2.7,
    inherit.aes = F) +
  scale_x_discrete("Block") +
  scale_y_continuous("PSE (ms VOT)") +
  scale_colour_manual(
    "Condition",
    labels = c("baseline", "+10ms", "+40ms", "prior"),
    values = c(colours.condition, "darkgray"),
    aesthetics = "color") +
  coord_cartesian(xlim = c(0.75, 7), clip = "off") +
  theme(
    axis.text.x = element_text(angle = 22.5, hjust = 1),
    plot.margin = unit(c(1, 3.2, 1, 1), "lines"),
    legend.position = "none")
```


But neither error-based learning nor Bayesian belief-updating models in their current forms, are sufficient to explain the marked slowdown and lack of convergence on the ideal boundary after the final exposure block.  
Under the conceptualisation of ideal adaptor models [@kleinschmidt-jaeger2015; @xie2023] the pace of adaptation or belief-updating is a function of confidence in prior expected values of the means and variances of the phonetic categories. The rapid change seen after the first exposure block indicates listeners have low confidence in their prior beliefs about the distributions and are therefore quick to adjust their categorisations---the shifted boundary should be dominated by the observations from the recent input rather than their prior beliefs. But weak prior beliefs under such a model of adaptation implicates sustained rapid shifts as the listener receives more evidence from the input which is not the result we found. For the +40 condition we even saw a retraction of the estimated boundary in test block 4. It is unclear whether this indicates that listeners had completely stopped adapting or if further shifts would have been observed with more exposure input.



Two previous studies have reported related results. @kleinschmidt-jaeger2016 exposed five different groups of listeners to VOT distributions for /b/ and /p/ that were shifted to different degrees. The five different exposure conditions were each shifted by 10msecs in VOT relative to the other, but held constant the distance between the /b/ and /p/ mean (always 40ms) and the variance of /b/ and /p/ (both always 8.3ms$^2$). All listeners were exposed to 222 trials of exposure input. Unlike the present study, @kleinschmidt-jaeger2016 did not include a pre-test or incremental intermittent testing. Instead, the effect of exposure was estimated by estimating listeners' categorization functions over the last third of the 222 trials. This revealed adaptation patterns that resemble those of the present study in several ways. First, the further the exposure distributions were shifted relative to the distributions expected from a 'typical' talker, the more listeners' categorization function was shifted in the same direction. Second, when expressed as a proportion of the shift expected from an idealized learner who has fully converged against the exposure distributions, larger shifts in the exposure distribution yielded *proportionally smaller* shifts in listeners' categorization functions ('shrinkage'). Third and finally, shrinkage effects were more pronounced for leftward- than rightward-shifted exposure distribution [see re-analysis in @kleinschmidt2020].^[@kleinschmidt2020 also presents additional experiments with extreme examples of leftward-shifted exposures. In his Experiment 4, different groups of listeners were exposed to /b/-/p/ distributions for which the /b/ mean was shifted to -20ms, -50ms, or even -80ms, while the /p/ mean remained at 50ms VOT. This experiment replicated the strong shrinkage effects for such leftward shifts: even in the most extreme condition (-80ms), listeners' average PSE was still larger than 20ms. These results should, however, be interpreted with caution since the experiment returned small shifts even for rightward shifts---possibly due to data quality issues with Mechanical Turk.]

Unlike the present study, @kleinschmidt-jaeger2016 and @kleinschmidt2020 used synthesized stimuli that sounded robotic, and were not designed to exhibit natural correlations between VOT, f0, and vowel duration. The distribution that listeners were exposed were *not* randomly sampled but rather were perfectly symmetric around the category means, and had identical variance for /b/ and /p/. This follows other DL studies [e.g., @clayards2008; @theodore-monto2019] but deviates from the distributions listeners are familiar with from natural speech. The fact that the present study resulted in 'shrinkage' effects similar to those found in @kleinschmidt2020 suggests that these effects are not an artifact of the lower ecological validity of the stimuli used in previous studies.





Our results replicate those general findings. Among the left-shifted conditions, the further the exposure distribution from the prior, the lower the proportion of boundary shift. The ideal listener boundaries for the baseline and +40 conditions particularly were almost equidistant from the prior in opposite directions (+21ms vs. -18ms relative to the prior) but boundary changes in the baseline condition vascillated within a very tight range, achieving at most about a fifth of the total distance which was half the proportion achieved by the +40 condition. This asymmetry is best explained when taking into account listener prior beliefs. That listeners in the +40 conditioned moved their boundaries more freely corresponds to the distribution of /t/ tokens in encountered speech. Figure \@ref(fig:exposure-means-database-matrix-plot) shows the widely distributed VOT values for /t/ across two different speech styles indicating that listeners are likely to have encountered a wide range of /t/ tokens in their lifetime. In contrast, the distribution of /d/ tokens is much more narrow in variability. The asymmetry in the distribution of /t/ and /d/ tokens in the input is likely to have resulted in listeners having stronger constraints on movements of the left-shifted conditions since adapting would mean interpreting more /d/s as /t/s under prior expectations that /d/s are much more tightly clustered in the cue space. Furthermore, as the voiced category is shifted further leftwards more negative VOT tokens will be included in exposure. Pre-voicing in L1-US English while quite common, is not a primary cue for identifying the voiced category. While it could reinforce perception of a voiced category research on pre-voicing effects in true-voicing languages suggest it is unlikely to elicit gradient responses in the perception of the voiced category [@vanalphen-mcqueen2006; @vanalphen-smits2004; @clare-schertz2022]. This means that for the voiced category as VOT gets smaller and moves into the negative region there is less information to be gained from attending to the changes. This flip-side to the effect on perception as VOT increases makes sense because there is not a category to the left of the voiced one in a dual-contrast language (it's /d/s all the way down).

Constrained adaptation should not be surprising since the need for perception to strike a balance between stability and flexibility has long been recognised. Constrained adaption is a logical result of a rational perceiver.

## Limitations
confirmatory. future work should compare other models [@xie2023]

stronger tests of potential constrain: run for longer

scaling ecological validity.

# Conclusions
...

These insights were facilitated by a incremental exposure-test paradigm and a single Bayesian psychometric mixed-effects analysis. Psychometric models have long been used in research on, for example, visual decision-making [e.g., @klein2001; @prins2011; @wichmann-hill2001]. For the advantages of Bayesian psychometric models, we refer to @kuss2005 and @prins2019. Here, we used a mixed-effects version of such a model in order to fit a single model across all participants, while correcting for participant-specific lapse rates and while modeling block-by-block changes in participants' psychometric functions. While such models used to require expensive software, they can now be fit in freely available software [\texttt{R} @R-base; \texttt{rstan}, @R-stan; @R-stanarm; \texttt{brms} @R-brms_a]. We hope that the present study will encourage other researchers to use similar models in their own research.
