<!-- Do NOT knit this document. It is part of a larger document. Instead knit the main document (my-apa-formatted-article) -->

<!-- ^[A related line of work has used distributional learning or explicit training paradigms to study the acquisition of *novel* sound contrasts [e.g., @maye2002; @mccandliss2002; @pajak-levy2012; @pisoni1982]. These studies, too, have observed outcomes predicted by distributional learning models [for review, see @pajak2016].] -->

# General discussion
Over the last 20+ years, landmark studies in adaptive speech perception have demonstrated that listeners' interpretation of speech is not static. It can change with recent exposure, accommodating differences in pronunciation across talkers [for reviews, see @bent-baeseberk2021; @schertz-clare2020]. Research on accent adaptation [AA, @eisner2013; @schertz2016; @xie2017], perceptual learning [visually-guided perceptual learning (VGPL)/LGPL, @eisner-mcqueen2005; @kraljic-samuel2006; @kurumada2018; @norris2003; @reinisch-holt2014; @vroomen2007], and distributional learning [DL, @bejjanki2011; @idemaru-holt2020; @kleinschmidt2020; @nixon2016; @theodore-monto2019] suggests that this flexibility is achieved through changes in listeners' *categorization functions*---the mapping from acoustic or phonetic cues to the phonological categories that form the input to spoken language understanding: after exposure to an unfamiliar talker, listeners interpret physically identical speech input differently. 

Here, we have responded to recent calls to better characterize *how* these changes in listeners' categorization functions come about. We set out to test several basic predictions of distributional learning models. Distributional learning models implement the hypothesis that listeners learn the statistics of talkers' speech, and use this implicit knowledge to interpret subsequent utterances by the same talker [@clayards2008; @idemaru-holt2011; @kleinschmidt-jaeger2015; @mcmurray-jongman2011]. To this end, we modified a distributional learning paradigm to test four predictions about the incremental unfolding of adaptation.

We found that listeners' categorization functions changed incrementally with exposure. The direction and magnitude of that change depended on listeners' prior expectations based on previously experienced speech input from other talkers (prediction 1 - *prior expectations*), and both the amount and distribution of phonetic evidence in the exposure input from the unfamiliar talker (predictions 2a and 2b - *exposure amount & distributions*, respectively). The Bayesian hypothesis tests we conducted also suggest properties of adaptive speech perception that go beyond these qualitative predictions. First, we found that participants' categorization functions changed quickly with exposure, and that the speed of these changes slowed-down with additional exposure (prediction 3 - *diminishing returns*). Second, we found evidence of potential constraints on listeners' adaptivity, as well as asymmetries in these constraints depending on whether exposure distributions were shifted downwards or upwards on the VOT continuum---contrary to prediction 4 (*learn to convergence*).

These findings inform theory by characterizing the computational properties of the mechanisms underlying listeners' adaptivity. We begin a discussion of predictions (1) and (2a,b). Then we turn to questions about the rate of changes (prediction 3), before considering potential constraints on the initial moments of adaptive speech perception (prediction 4). We close by considering limitations of the present study and how future work can overcome them.

## Incremental adaptation based on the amount and distribution of phonetic evidence (Predictions 1 and 2a,b)
To the best of our knowledge, the present study is the first to assess how the joint effects of prior and recent exposure unfold gradiently with increasing exposure, testing predictions (1) and (2a,b). While most contemporary theories of adaptive speech perception share these qualitative predictions, few experiments have investigated how listeners' categorization functions change with continued exposure to a phonetic distribution. Next, we discuss notable exceptions to this trend, and how our results relate to those works.

<!-- We used a incremental exposure-test design to assess incremental changes in listeners' categorization functions from pre-exposure onward. We found that that listeners' categorization function shifted with increasing exposure to shifted phonetic distributions. The direction and magnitude of these shifts developed gradiently and were qualitatively consistent with the predictions of distributional learning accounts [e.g., @apfelbaum-mcmurray2015; @harmon2019; @johnson1997; @kleinschmidt-jaeger2015; @magnuson2020; @nearey-assmann2007; @sohoglu-davis2016; @xie2023]. This was the case both for shifts in listeners' categorization function relative to other exposure conditions (Table \@ref(tab:hypothesis-table-simple-effects-condition)) and for shifts in listeners' categorization function relative to their pre-exposure categorization function (Table \@ref(tab:hypothesis-table-simple-effects-block)). In both cases, every single test we conducted went in the direction predicted by distributional learning theories. We elaborate on these findings. -->

### Prediction 1: Adaptation begins with, and integrates, listeners' prior experience
The inclusion of a pre-exposure test in our design made it possible to assess prediction 1---that shifts in listeners' categorization function should depend on how the exposure distributions *relative to listeners' relevant prior experiences*. In line with this prediction, listeners' responses during the pre-exposure test were well approximated based on a database of /d/ and /t/ productions. While these effects of prior knowledge are often assumed, the present experiment is---to our knowledge---the first time they have been demonstrated for adaptive speech perception. 

The inclusion of a pre-exposure test also made it possible to test how prior expectations (prediction 1) and exposure inputs (prediction 2a,b) *jointly* explained the direction and magnitude of changes in listeners' categorization functions. We found that the direction of changes in listeners' PSEs *relative to pre-exposure* was predicted by the direction of the shift in /d/ and /t/ distributions relative to their distributions in prior experience (Figure \@ref(fig:prior-distributions-and-exposure-conditions)A). This joint effect of prior expectations and exposure input is predicted by distributional learning accounts that explain adaptive speech perception as incremental integration of listeners' prior expectations---based on previously experienced speech input---and the statistics of the exposure input. This includes rational theories of adaptive speech perception [e.g., the ideal adaptor framework, @kleinschmidt-jaeger2015; @kleinschmidt-jaeger2016], some theories of normalization [e.g., the probabilistic sliding template model, @nearey-assmann2007], as well as episodic [@goldinger1998] and exemplar models [@johnson1997]. Other distributional learning accounts are in principle compatible with our finding but would need to be expanded to incorporate prior expectations and the processes that integrate those expectations with new exposure input [e.g., @bejjanki2011; @mcmurray-jongman2011; see discussion in @persson2024; @xie2023].

Future tests will be necessary to replicate these effects of prior experience. Here we used Chodroff and Wilson's (2018) database of isolated speech because its recordings resemble the speech style of our stimuli (isolated productions with slower speech rate), as well as the gender, age range, and related acoustic characteristics (f0) of the talker we recorded our stimuli from. That is, we followed the common assumptions that listeners base their prior expectations for difference in the realization of phonetic cues due to linguistic context [e.g., speech rate, in line with @allen-miller1999; @miller-dexter1988; @utman1998] and general acoustic talker characteristics [e.g., f0]---either based on pre-linguistic cue normalization or based on rich storage of exemplars [for relevant discussion, see @apfelbaum-mcmurray2015; @baeseberk2014; @dilley-pitt2010; @johnson1997; @johnson-sjerps2021; @toscano-mcmurray2012]. The fact that this yielded a good fit to participants' responses during the pre-exposure test suggests that our assumption at least did no harm. Additional analyses not reported here---but easily replicable with the data and code provided on OSF---confirmed that an idealized pre-exposure model based on *connected* speech does not fit participants' responses as well (though all qualitative results remain the same).

Still, the database we used contained only slightly more than 1000 observations from thirteen female talkers. We aimed to remedy this downside by using five-fold cross-validation (see SI \@ref(sec:idealized-prior-listeners)), in order to quantify the uncertainty about the true range of a typical L1-US English listeners' prior expectations (the gray ribbons in Figure \@ref(fig:plot-fit-PSE)C). This does not, however, remove the need to validate our results based on new participants and larger phonetic databases.

### Prediction 2a: Adaptation increases with the amount of exposure
Our results also support prediction 2a---that the magnitude of changes in listeners' categorization function should gradiently increase with the *amount* of exposure. This is another prediction that is often assumed but rarely tested. Here, it received support from the comparisons across blocks: increasing exposure consistently yielded additional shifts in listeners' PSE. In particular, the Latin-square design over exposure blocks completely de-correlated the amount of exposure that participants in a given exposure condition received from the specific distribution of phonetic properties (cf. Figure \@ref(fig:design-distribution)). This contrasts with existing LGPL studies, in which exposure amount was confounded with differences in exposure distributions [@cummings-theodore2023; @liu-jaeger2018; @liu-jaeger2019], and VGPL studies in which exposure was limited to many repetitions of a single stimulus [e.g., @kleinschmidt-jaeger2012; @vroomen2007].

In VGPL/LGPL paradigms, listeners are exposed to natural recordings of one phonetic category (e.g., /s/) and shifted instances of a second category that are manipulated to be perceptually more similar to the first category (e.g., /s/-like /`r linguisticsdown::cond_cmpl("ʃ")`/). Both the typical and the shifted sound instances are lexically or visually labeled by their context. For example, in an LGPL study, the lexical context will disambiguate the intended category of both the typical sounds (e.g., "dino*s*aur") and the shifted sounds [e.g., "medi*sh*ine", @norris2003; @eisner-mcqueen2005; @kraljic-samuel2005]. <!-- Studies like this typically compare two groups of listeners that differ only in which of the two sounds was shifted. For example, one group of listeners might be exposed to 20 typical /s/ and 20 shifted /s/-like /`r linguisticsdown::cond_cmpl("ʃ")`/, mixed with 160 filler words that do not contain either sound. The other group of listeners might be exposed to 20 typical /`r linguisticsdown::cond_cmpl("ʃ")`/ and 20 shifted /`r linguisticsdown::cond_cmpl("ʃ")`/-like /s/, mixed with the same 160 filler words. Following exposure, the two groups of listeners will categorize sounds along an unlabeled test continuum (e.g., "asi" to "ashi") differently. Specifically, listeners in each group will categorize more sounds along the continuum as belonging to the category that was shifted during exposure [e.g., @norris2003; @eisner-mcqueen2005; @kraljic-samuel2005]. -->
In a particularly informative study, @cummings-theodore2023 compared shifts in categorization functions between groups of listeners after exposure to 1, 4, 10, or 20 lexically labeled shifted /s/ or /`r linguisticsdown::cond_cmpl("ʃ")`/ tokens (each matched by an equal number of unshifted tokens from the opposite category). Shifts in listeners' categorization functions increased with the number of exposure to tokens, in line with prediction 2a of distributional learning models. @vroomen2007 found similarly increasing shifts in categorization functions *within* participants, comparing the effects of 1, 2, 4, ..., 32 exposures to visually labeled shifted tokens [see also @kleinschmidt-jaeger2012].^[With increasing exposure, the direction of shift begins to reverse [returning to baseline after 128-256 exposures, @kleinschmidt-jaeger2011; @vroomen2007] and can even change directional altogether, depending on the degree of the shift [@kleinschmidt-jaeger2012]. Later work showed that distributional learning models predict this reversal due to the specific choice of stimuli used in those experiments <!-- , which attribute it to the fact that the VGPL paradigm presents the exact same stimulus on each exposure trial --> [@kleinschmidt-jaeger2015].] 

The present study demonstrated similarly gradient effects with increasing exposure to a mixture of labeled and unlabeled exposure tokens that were randomly sampled from a *distribution* of phonetic tokens, more closely resembling the situation listeners would experience in everyday speech perception. This adds to a growing number of similarities in the findings between LGPL/VGPL and DL paradigms, as expected under the hypothesis that changes observed in both paradigms originate in the same underlying mechanisms [see discussions in @kleinschmidt2015; @zheng-samuel2020; @xie2023].

### Prediction 2b: Adaptation depends on the phonetic distribution in the exposure input
Prediction 2b---that the direction and magnitude of changes in listeners' categorization function should depend on the *phonetic distribution* of the exposure input---is perhaps the best documented prediction of the ones we tested [@bejjanki2011; @chladkova2017; @clayards2008; @colby2018;  @kleinschmidt-jaeger2011; @kleinschmidt-jaeger2012; @kleinschmidt-jaeger2016; @nixon2016; @saltzman-myers2021; @theodore-monto2019]. However, previous work has exclusively assessed this prediction by comparing the outcome of adaptation between exposure conditions.

For example, @kleinschmidt-jaeger2016 exposed five different groups of listeners to VOT distributions for /b/ and /p/ that were shifted to different degrees. The five different exposure conditions were each shifted by 10ms in VOT relative to the other, but held constant the distance between the /b/ and /p/ mean (always 40ms) and the variance of /b/ and /p/ (both always 8.3ms$^2$). All groups of listeners were exposed to 222 trials of exposure input. As is the norm for DL experiments, Kleinschmidt and Jaeger did not include a pre-test or incremental intermittent testing. Instead, the effect of exposure was evaluated by estimating listeners' categorization functions over the last third of the 222 trials [another common approach is to average over *all* trials, e.g., @clayards2008; @nixon2016]. This revealed that listeners' categorization functions differed between exposure conditions in ways consistent with distributional learning models: the more the exposure distribution was shifted rightwards relative to each other, the more listeners' categorization functions also were shifted in the same direction.

The present work extends these findings in three ways. First, as already mentioned above, it is the first demonstration that both prior experience and the exposure input need to be taken into account in order to correctly predict changes in listeners' categorization function. Second, we demonstrated gradient *incremental* adaptation towards the exposure distribution. We found that the direction of the shift of the /d/ and /t/ category means in the exposure input correctly predicted the relative ordering of listeners' PSEs in Test 1-4. We also found that shifts in category means of larger magnitude (+20ms vs. -20ms compared to -10ms vs. -20ms) yielded larger absolute shifts in listeners' PSE. 

Third and finally, these changes in listeners' categorization functions were observed for natural-sounding speech that followed the types of heterogeneous distributions of phonetic cues listeners would be likely to experience during everyday speech perception. The use of natural-sounding stimuli contrasts with the type of clearly robotic sounding speech used in many previous DL experiments on VOT [for notable exceptions, see @chladkova2017; @theodore-monto2019]. Robotic speech which might lead listeners to adopt different strategies than they would normally use. In particular, such speech provides a clear signal to listeners that some of their expectations about typical speech inputs are unlikely to extend to the current input, which might inflate listeners' readiness to adapt their perception. Similarly, our use of naturalistic *phonetic distributions* contrasts with the use of exposure distributions that are perfectly symmetric around their mean, or that in other ways differ from the distributions listeners experience in real life [for a notable exception, @chladkova2017]. 

## Prediction 3: Rapid adaptation with *diminishing returns*
Having established that exposure led to gradient changes in participants' categorization behavior, we turn to the rate with which these changes unfolded with additional exposure. The rate of change in listeners behavior is of theoretical interest for two reasons. First, it speaks to the plausibility that the same mechanisms that drive adaptive behavior in the present DL paradigm also underlie adaptive behavior during everyday speech perception (which has been found to be *very* fast, as we discuss below). Second, it speaks to the nature of the learning mechanisms that underlie adaptive speech perception. 

### How quickly can listeners adapt their speech perception?
We found significant shifts in listeners' categorization function even after the briefest exposure tested. Exposure to 24 tokens each of shifted /d/ and /t/ was sufficient to significantly change how listeners interpreted subsequent inputs. Of note, only half of these exposure tokens labeled the intended category, the other half did not. Even when trials were labeled, labeling was indirect rather than through explicit feedback: on labeled trials, the two response options listeners saw both had the same onset stop (e.g., "din" and "dill"). Previous DL studies have assessed exposure effects after *much* longer exposures, typically hundreds of trials [e.g., 192 trials in @harmon2019; 200 in @idemaru-holt2011; 222 in @clayards2008; @kleinschmidt-jaeger2016; 236 in @theodore-monto2019; 456 in @nixon2016]. The present results demonstrate that a fraction of the amount of exposure employed in previous studies is sufficient to elicit changes in listeners' categorization behavior.

This finding informs ongoing discussions that the type of adaptive changes observed in AA and LGPL/VGPL paradigms could plausibly arise from the same mechanisms as those observed in DL paradigms like in the present study [@bradlow-bent2008; @baeseberk2018; @zheng-samuel2020; @xie2023]. In the introduction, we mentioned findings of improved speed and accuracy of cross-modal priming after exposure to only 18 sentences from an L2-accented talker---the shortest tested exposure we are aware of [@clarke-garrett2004; @xie2018]. Other AA work has more directly demonstrated that exposure changes listeners' categorization functions. For example, @xie2017 found changes in listeners categorization behavior after exposure to only 30 critical second language-accented words. The direction of these changes was consistent with distributional learning accounts of adaptive speech perception. Together with evidence from additional experiments, Xie and colleagues concluded that "listeners dynamically update their own cue-weighting functions during rapid phonetic adaptation to foreign accents, and critically, over much shorter time span[s] than shown in previous studies of second language phoneme learning" (p. 215). By demonstrating that DL paradigms can elicit qualitatively similar changes with similarly little exposure, the present study lends further plausibility to the hypothesis that these changes are driven by the same underlying mechanisms.

Some experiments on LGPL/VGPL have demonstrated effects after even less exposure, with detectable changes in listeners' categorization responses after as few as 2-4 exposures to visually or lexically labeled phonetically tokens [@cummings-theodore2023; @kleinschmidt-jaeger2011; @kleinschmidt-jaeger2012; @liu-jaeger2018; @liu-jaeger2019; @vroomen2007]. In comparing findings across paradigms, future work should keep in mind that DL and LGPL/VGPL paradigms differ in the amount of information conveyed by each exposure token. LGPL/VGPL paradigms typically employ exposure stimuli that are a) labeled and b) auditorily maximally ambiguous---falling between the two categories that the experiment focuses on. Distributional learning accounts predict that such stimuli should be highly informative, leading to comparatively large changes in categorization behavior. This is in line with recent findings: when stimuli in LGPL/VGPL experiments are shifted less than to the point of maximal auditory ambiguity, listeners exhibit smaller shifts in categorization behavior [@babel2019; @kleinschmidt-jaeger2012; @tzeng2021; see also @cummings-theodore2023]. 

In contrast, DL paradigms a) typically employ only unlabeled stimuli [@clayards2008] or a mixture of unlabeled and labeled stimuli [e.g., @kleinschmidt-jaeger2016 and the present paradigm], and b) reflect a *distribution* of phonetic properties---ranging from more to less expected under listeners' prior expectations. This makes the speech inputs in DL paradigms more similar to what one would expect during exposure to natural accents and other cross-talker differences. But it also means that exposure tokens in DL experiments are, on average, considerably less informative than in an LGPL/VGPL experiment. Future work that aims to compare the speed of adaptive speech perception across these two paradigms should thus do so *relative to the amount of information conveyed by each exposure*.

<!-- TO DO: *could* add a figure here showing how the PSE would be predicted to change for our experiment under a certain kappa, nu combination, and compare that against PSE changes for the same data when all /d/ stimuli are set to *prior* PSE value and all /t/ stimuli are set to prior /t/ (or vice versa) -->

### First fast, then slow: *diminishing returns* of exposure
Our comparisons across test blocks within each exposure condition found suggestive---but not decisive---evidence that the speed of incremental changes in listeners' PSE decreased with increasing exposure: the same amount and distribution of phonetic evidence yielded smaller *additional* changes in listeners' PSE, the more exposure blocks listeners had already experienced.^[Here, we focus on changes in listeners' behavior during exposure to *stationary* statistics. Another line of work has shown listeners might become less sensitive to *changes* in a talker's speech statistics after prolonged exposure [e.g., @kraljic-samuel2011; @saltzman-myers2021; but see @theodore-monto2019]---a finding predicted if adaptive speech perception is an active process [@magnuson-nusbaum2007] that requires change detection [@qian2012].] To the best of our knowledge, this is the first study to report this pattern of gradiently diminishing returns. A similar pattern is, however, indirectly evident in at least one other recent study. Kleinschmidt [-@kleinschmidt2020, Experiment 3] re-analyzes several DL experiments originally presented in @kleinschmidt-jaeger2016. Since these experiments lacked incremental testing, Kleinschmidt entertained several approaches to estimating incremental exposure effects, while controlling for differences in the phonetic properties of the stimuli over which these effects were assessed. The result<!---summarized here in Figure \@ref(fig:kleinschmidt2020)--->is unsurprisingly much noisier than in the present study, but the pattern is compatible with the hypothesis that adaptation initially proceeded quickly, and then increasingly more slowly. Paralleling the present experiment, this is particularly evident for the more extreme exposure conditions. Finally, rapid initial changes with diminishing returns have also been observed in some AA studies [@bieber2023; @xie2021jep, SI $\S3$]. However, in these studies changes are often measured in terms of changes in the percentage of accurate responses---a bounded space in which improvements *inevitably* must become smaller as they approach ceiling performance. The present study avoids this issue.

<!--(ref:kleinschmidt-increment-replotted) Changes in listeners' PSE as a function of exposure for the experiments reported in @kleinschmidt-jaeger2016. Replotted with aesthetic changes. As in Figure \@ref(fig:plot-fit-PSE), point ranges show the 95% bootstrapped CIs over by-participant means. Following Figure 9 from @kleinschmidt2020, the 222 trials of each exposure condition were split into six bins of 37 trials each. -->

```{r kleinschmidt-increment-replotted, fig.cap="(ref:kleinschmidt-increment-replotted)"}
# TO DO: add description of Dave's study; incl. absence of incremental testing, and the way it was 'hacked'/approximated in the modeling in section XXX of his ms. NOTE: this is NOT about shrinkage. it's about *diminishing returns*
```

```{r prediction-error-during-exposure}
d.exposure_surprisal <-
  d.for_analysis %>%
  filter(Phase == "exposure", Item.Labeled) %>%
  cross_join(
    io %>%
      select(-x) %>%
      rename(IO.Type = Condition.Exposure)) %>%
  # For each labeled exposure stimulus, calculate the surprisal of seeing the category label
  # (this is not quite parallel to what we did for test tokens, which was looking at the
  # surprisal of the VOT, rather than the surprisal of the category label)
  mutate(
    x = ifelse(str_starts(IO.Type, "prior"), pmap(list(Item.VOT, Item.f0_Mel, Item.VowelDuration), ~ c(...)), Item.VOT),
    posterior = pmap_dbl(
      .l =
        list(
          x,
          category,
          io),
      .f = function(x, y, z)
        pmap(
          list(z$mu, z$Sigma, z$Sigma_noise),
          ~ dmvnorm(x, ..1, ..2 + ..3)) %>%
        reduce(~ ..1 / (..1 + ..2)) %>%
        { if (y == "/d/") . else 1 - .}),
    surprisal = -log2(posterior))

d.exposure_surprisal %<>%
  group_by(Condition.Exposure, IO.Type) %>%
  summarise(across(surprisal, list(mean = mean, sd = sd)))
```

Such *diminishing returns* of exposure are explicitly predicted only by some distributional learning models. This includes error-driven learning models [e.g., @davis-sohoglu2020; @harmon2019; @olejarczuk2018; @sohoglu-davis2016] and models of ideal information integration [ideal adaptors, @kleinschmidt-jaeger2015; @kleinschmidt-jaeger2016]. Prior to adaptation, an idealized listener (horizontal gray ribbons in Figure \@ref(fig:plot-fit-PSE)C) would, on average, experience larger prediction errors (mean surprisal per exposure input in -20ms condition ${\rm E}[-\log_2 p(category | VOT, f0, vowel\ duration)] =$ `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift0", str_starts(IO.Type, "prior")) %>% pull(surprisal_mean) %>% mean(.) %>% round(., 2)` bits; -10ms condition = `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift10", str_starts(IO.Type, "prior")) %>% pull(surprisal_mean) %>% mean(.) %>% round(., 2)` bits; +20ms condition = `r d.exposure_surprisal %>% filter(Condition.Exposure == "Shift40", str_starts(IO.Type, "prior")) %>% pull(surprisal_mean) %>% mean(.) %>% round(., 2)` bits). As listeners converge towards the distribution of /d/ and /t/ in the exposure condition, they should experience increasingly smaller prediction errors (or equivalently: less new information) processing the exposure tokens. An idealized learner that has fully converged against the exposure distributions (colored lines in Figure \@ref(fig:plot-fit-PSE)C) would, on average, experience only `r d.exposure_surprisal %>% filter(Condition.Exposure == IO.Type) %>% pull(surprisal_mean) %>% mean(.) %>% round(., 2)` bits of surprisal per exposure input.

Models of adaptive speech perception that predict adaptation to be a positive monotonic function of the prediction error, thus offer an explanation for the diminishing returns of exposure observed in the present study. Of note, they do so without introducing arbitrary changes in learning rates or other parameters: it is the decrease in additional information gained from additional exposure that drives the decreasing rate of change in listeners' behavior.<!-- TO DO: revisit after review whether we'd like to include this: ^[This also highlights that the observable behavior---e.g., changes in listeners' categorization responses---ought not to be confused with the mechanism itself. For example, a *lack* of changes in behavior does not *necessarily* imply a lack of learning (even beyond the usual caveats that apply to null effects). This would only follow if it is also shown that learning would be expected to lead to changes in behavior. But this would require a learning model, a linking hypothesis, and their application to the exposure input---something that is usually lacking from the interpretation of null effects [cf. @zheng-samuel2020].] --> If the pattern of diminishing returns is replicated in future work, this would raise questions as to whether similar predictions follow from other distributional learning accounts [e.g., C-CuRE normalization, @mcmurray-jongman2011; exemplar models, @johnson1997; DNNs, @magnuson2020]. If, on the other hand, future tests reliably fail to replicate these findings, this would constitute a serious challenge to models that predict adaptation to be sensitive to the prediction error.

## Prediction 4: Constraints on the early moments of adaptive speech perception? 
At first glance, it would be tempting to interpret the 'diminishing returns' discussed in the previous section as evidence that listeners have converged against the exposure distributions in the input---i.e., that adaptation has successfully completed. In particular, we found that there was at best anecdotal evidence that the final exposure block had *any* additional effect (cf. hypothesis tests in Table \@ref(tab:hypothesis-table-simple-effects-block)). However, the comparison against idealized learners (Figure \@ref(ref:)) revealed that listeners had *not* actually successfully learned the phonetic distributions of the unfamiliar talker. Rather, listeners appear to have converged prematurely, long before they achieved the theoretically possible recognition accuracy. 

As we discuss next, premature convergence---if confirmed---would introduce an important new constraint that theories of speech perception need to explain. To the best of our knowledge, no existing theory has made this specific prediction. Previous literature does, however, offer a distinction between *model learning* and *model selection* that might point the way to an explanation. We first elaborate on this point, as it helps to establish the theoretical relevance of this finding. 

### Why is premature convergence theoretically relevant?
One explanation for premature convergence is that adaptive speech perception---or at least its earliest moments---is *not* the result of distributional learning. This would require the formulation of, as of yet unspecified, alternative mechanisms. It would also require explanations for the properties of adaptive speech perception that *do* receive explanations under the hypothesis that adaptive speech perception involves distributional learning, including the properties we have demonstrated above.

There are, however, alternative explanations that do not throw out the proverbial baby with the bathwater. Under the first of these explanations, adaptive speech perception *is* driven by distributional learning, just not *unconstrained* distributional learning. This is the hypothesis we consider most plausible at this point (while maintaining considerable uncertainty about it), and so we elaborate on it in some detail. Specifically, it is possible that rapid adaptation during the earliest moments of encountering an unfamiliar talker is limited to the selection of (mixtures of) previously experienced talkers. 

To appreciate this idea, consider the distinction between *model learning* and *model selection*, proposed by Xie and colleagues [-@xie2018, p. 2028-2029]. Model learning refers to listeners learning new phonetic category representations for the unfamiliar talker. For example, proposals that listeners store speech episodes [@goldinger1998; @goldinger2007] or exemplars [@hay2018; @johnson1997] from the unfamiliar talker, and then use these to categorize subsequent speech from that talker constitute a form of model learning. Critically, as such learning continues, listeners should increasingly come to reflect categorization behavior that is based on the phonetic distributions of the new talker---contrary to what we seem to observe. Similarly, unconstrained ideal information integration would constitute model learning [as implemented in the ideal adaptor model, @kleinschmidt-jaeger2015].

Model selection, on the other hand, refers to the idea that listeners select between different *previously experienced* models. In the case of speech perception, the models being selected between are talker- or talker group-specific phonetic representations [e.g., idiolects, dialects, sociolects, etc., as reviewed in @pajak2016]. Each of these models specifies a mapping from phonetic cues to categories. Model selection, too, can be seen as a form of distributional learning, as the incremental reweighting of different models based on both top-down (contextual) and bottom-up (acoustic) cues to talker identity [@kleinschmidt2015, p. 180-182]. This reweighting allows listeners to adapt to unfamiliar input by upweighting previously learned models that more accurately categorize speech from the new talker. Unlike model learning, however, the flexibility afforded by model selection is limited, and very strongly constrained by previous experience. Specifically, model selection alone would only allow listeners to adapt their behavior *up to the most extreme* previously stored model. This is why Kleinschmidt and Jaeger hypothesized that human speech adaptation might draw on both model selection and model learning in order to strike a trade-off between stability and flexibility [@kleinschmidt2015, Equations 24-25, p. 181]. But what if changes in listener's behavior are primarily determined by model selection, at least during the earliest moments of exposure? @xie2018 argued that there are reasons to believe this to be the case, and that model learning might depend on slower neural mechanisms, such as memory consolidation of new exemplars during sleep [for relevant discussion, see @estes1986; @fenn2013; @xie2018sleep]. 

## Limitations and future directions
The present study set out to investigate incremental adaptation to a single talker, whose pronunciations were shifted relative to listeners' prior expectations. The conclusions we have discussed so far should be interpreted in light of several limitations of the approach we took. First, our experiment investigated incremental adaptation to a single female talker's productions of syllable-initial /d/-/t/ by L1-US English listeners. In particular, our exposure conditions shifted the distribution of VOT and, by extension, f0 and vowel duration. As adaptive speech perception---including its generalization across categories, phonetic contexts, and talkers---can depend on the exposure talker, the phonetic contrast or the phonetic cues it involves [e.g., @eisner-mcqueen2005; @kraljic-samuel2007; @mitterer2013; @xie2017; @xie2021jep], future work is necessary to assess how general the present findings are. For instance, while some phonetic contrasts rely on features that vary primarily between talkers, other contrasts vary primarily within talkers (e.g., depending on changes in speech rate or style). If listeners have implicit knowledge of such differences, this might affect the speed or even the convergence behavior for those types of phonetic contrasts [cf. discussion in @kleinschmidt-jaeger2015; @kraljic-samuel2007].

Second, the present study shares with other DL paradigms that a comparatively small number of minimal pair items was repeated many times (though this repetition was substantially reduced in the present work, compared to previous work, with more than 50 distinct stimuli for the 144 trials), with only minimal phonetic differences embedded in otherwise constant phonetic contexts (e.g., the vowel following /d/-/t/ was always the same), and presented in isolation. This sacrifice of ecological validity was motivated by our goal to test stronger predictions about the direction and relative magnitude of effects (rather than merely the existence of effects). It does, however, mean that the speech input that participants experienced in the experiment differed from everyday encounters with unfamiliar talkers: listeners typically experience *connected* speech from unfamiliar talkers, which tends to be produced with faster speech rates and comes with additional segmentation challenges; while the same phonetic contrast might appear many times, it will not necessarily appear in the same phonetic context, least of all in the same word; and the speech of talkers with unfamiliar accents often deviate from listeners' expectations in more than a single segmental contrast. Comparatively little is known about adaptive speech perception under such more common conditions [even AA studies have typically focused on short isolated sentences, @bradlow-bent2008; @clarke-garrett2004; @sidaras2009, and the many studies inspired by these pioneering works]. <!-- It is possible, for example, the rapidity of adaptive changes observed in the present study over-estimates the adaptivity of everyday speech perception. There is evidence, for example, that the repetition of minimal pair recordings can affect listeners categorization responses [@lancia-winter2013], perhaps because it helps listeners isolate relevant differences between recordings.-->

Third, as already mentioned, some of the tests were conducted post-hoc, and thus should be interpreted with caution. In particular, future experiments with longer exposure would provide more decisive tests of the explanations we offered for premature convergence. Additional data from future applications of the incremental exposure-test paradigm with different phonetic shifts will also facilitate stronger quantitative tests of models of adaptive speech perception [in the spirit of @coretta2023; @guest-martin2021; @yarkoni-westfall2017; @xie2023]. In a recent review of the field, @xie2023 demonstrated that the signature findings of some of the most popular paradigms in adaptive speech perception do not distinguish between radically different theoretical accounts. Qualitative improvements in speech recognition can be explained by mechanisms ranging from early pre-linguistic perceptual normalization, changes in the representations of phonetic categories, or upstream changes in decision-making. Xie and colleagues concluded that the effective comparisons of these theories will require quantitative data sets that constrain the way in which listeners' categorization behavior changes depending on the amount and nature of the input. The use of incremental testing and multiple exposure conditions with different phonetic shifts---as explored in the present study---provides such data [see also @xie-kurumada2024].

<!-- A second reason for the relative scarcity of research on incremental changes in speech perception might be that repeated testing comes with its own unique challenges. For example, if test tokens are sampled from natural accents---the most common approach in AA research---these tokens can themselves contain information about the target accent, thereby masking exposure effects [for a particularly clear demonstration, see @xie-kurumada2024]. Even for paradigms that employ synthesized stimuli, or otherwise control the informativity of test tokens [e.g., @chodroff-wilson2020], repeated testing can come with challenges. As already discussed, there is now evidence that repeated testing over unlabeled uniform test continua can reduce the effects of exposure [@scharenborg-janse2013; @cummings-theodore2023; @zheng-samuel2023; @liu-jaeger2018; @liu-jaeger2019; @giovannone-theodore2021; @tzeng2021]. This is compatible with some distributional learning accounts of adaptive speech perception [for discussion, see @kleinschmidt-jaeger2015]: if the same learning mechanisms that operate during labeled and unlabeled exposure trials continue to operate during unlabeled test trials, the unexpected uniform distribution over the VOT continuum during test blocks should begin to undo the effects of preceding exposure.^[The specific predictions depend on the---as of yet unknown---way listeners adapt to unlabeled speech inputs [for an initial comparisons of several candidate models for unsupervised adaptation, see @yan-jaeger2018]. Additionally, it is possible that adaptation to unlabeled inputs involves different mechanisms than adaptation to labeled trials.] In the present study, we replicated this effect of repeated testing for the final three test blocks. -->

# Conclusions
Research on adaptive changes in speech perception has made great strides since the foundational works more than twenty years ago. Now that the existence of adaptive changes in speech perception is no longer in question, recent reviews of the field have emphasized the need to develop novel paradigms that can inform the functional relation between exposure inputs and changes in listeners' perception [@coretta2023; @schertz-clare2020; @xie2023]. The present study is a response to this call. We set out to more clearly characterize the incremental unfolding of adaptation to changes in the realization of a simple two-way phonetic contrast (/d/-/t/). This allowed us to assess several predictions of distributional learning accounts of adaptive speech perception, including previously untested predictions. 

In line with these theories, we found that listeners initially draw on prior experience with other talkers to recognize input from the unfamiliar talker. With increasing exposure, listeners then adapted their categorization responses, improving recognition accuracy. The incremental unfolding of these changes followed the prediction of distributional learning accounts---in particular, accounts that predict changes in listeners' perception to depend on the prediction error (or the amount of new information) associated with each new exposure input. Finally, we found suggestive evidence that adaptivity, at least during the earliest moments of exposure, is constrained in ways that do not seem to be predicted by existing models: while adaptation was rapid, it also slowed down and seemed to converge against a stable state long before listeners approached the recognition accuracy expected from an idealized learner. The specific nature of these constraints strikes us as an important target for future work, as they potentially impose novel constraints on theories of adaptive speech perception.

Our findings were facilitated by both Bayesian psychometric mixed-effects analysis [e.g., @prins2019bayesian; @kuss2005] and normative models of adaptive speech perception [@kleinschmidt-jaeger2015]. We extended the former to fit a single model across all participants, while correcting for participant-specific lapse rates and while modeling block-by-block changes in participants' psychometric functions. While such models used to require expensive software, freely available software has substantially lowered the entry cost for such approaches [e.g., \texttt{R}, @R-base; \texttt{brms}, @R-brms_a]. Similarly, there are now R libraries that facilitate the fitting and evaluation of ideal observers and adaptors [\texttt{MVBeliefUpdatr}, @R-MVBeliefUpdatr; \texttt{beliefupdatr}, @beliefupdatr]. The R markdown available on OSF provides a starting point for other researchers interested in either type of model.

```{r}
rm(fit_test.upto.test4)
```

